### 2016-10-05 titan.tex
titan text

\subsection{OLCF and Titan}

In 2004, the Oak Ridge Leadership Computing Facility (OLCF) was established at
Oak Ridge National Laboratory (ORNL).The  OLCF's is mission to accelerate the
scientific discovery and engineering progress by providing outstanding
computing and data management resources to high-priority research and
development projects.

Titan, a leadership-class and the fastest open-science supercomputer in the
U.S., is located at OLCF. Titan enables scientists to evaluate and assess various
complex physical phenomena via large-scale computational simulations.

Titan is a Cray XK7 system with 18,688 compute nodes~\cite{top500}. Each
compute node is equipped with an AMD Opteron CPU and a Nvidia Tesla GPU. This
hybrid design provides improved energy efficiency, as well as an order of
magnitude in computational capacity over its predecessor. Titan has 710 TB of
total system memory. This leadership capability compute system is backed by a
center-wide parallel file system known as Spider II~\cite{spider2}. Spider II
is one of the world's fastest and largest POSIX complaint parallel file
systems. It is designed to serve write-heavy I/O workloads generated by Titan
compute clients and other OLCF resources.

The topology and architecture details of Spider II infrastructure are
illustrated in Figure~\ref{fig:xk7-compute-node} and
described as follows:

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/spider2arch.pdf}}\\
    \end{tabular}
    \caption{Infrastructure and I/O path between Titan (Cray XK7) and its backend storage}
    \label{fig:xk7-compute-node}
\end{figure}

Spider II is based on the Lustre technology~\cite{lustre}.  It is configured
and deployed as two independent, non-overlapping file systems, each with 144
Lustre Object Storage Servers (OSSs) and 1,008 Lustre Object Storage Targets
(OSTs). Spider II offers ~30 PB usable file system capacity to OLCF systems and
users and has over 1 TB/s aggregate write I/O performance.