### 2017-03-19 WMSLiterature.tex
First pass section II; moved definitions to end of introduction

The Production and Distributed Analysis (PanDA) is a Workload Management System
(WMS) developed to support the operations of the LHC ATLAS experiment. PanDA is
specifically design to manage the execution of distributed workloads and
workflows via pilots.

WMS for distributed executions are a type of middlware, designed to coordinate
specific activities about resources and workloads: discovering and selecting
resources, submitting the tasks of a workload, and monitoring the execution of
those tasks~\cite{marco2009glite}. Pilot is an abstraction that enables
multi-level scheduling by decoupling resource acquisition from workload
scheduling~\cite{turilli2015comprehensive}. Pilots are implemented by resource
placeholders: a request for a certain amount of resources is scheduled on a site
(e.g., by submitting a job or requesting a virtual machine) and once the
resources are acquired, tasks are scheduled directly to the pilot, not to site's
scheduler.

WMS like PanDA implement the pilot abstraction: WMS control the acquisition of
resources by means of pilots and then manage the execution of workloads or
workflows' tasks on those pilots. In tis way, pilots expose a uniform scheduling
interface and execution environment, abstracting the differences among the
heterogeneous resources of multiple sites; WMS centralize the management of the
execution process, offering a unified interface to the application layer.
Pilot-enabled WMS are particularly relevant for LHC experiments, where large
amount of heterogeneous resources, distributed across multiple sites have to be
coordinated to execute tens of million of tasks per months while producing and
storing petabytes of data.

Several pilot-enabled WMS were developed for the LHC experiments:
AliEn~\cite{Bagnasco2010} for ALICE; DIRAC~\cite{Paterson2010} for LHCb;
GlideinWMS~\cite{sfiligoi2008glideinwms} for CMS; and
PanDA~\cite{maeno2014evolution} for ATLAS. These systems implement similar
design and architectural principles: separation of concern between management
and execution, and data and compute; centralization of the management
capabilities; distributed execution of the execution capabilities across
multiple sites; unification of the application interface; hiding of resource
heterogeneity; centralized monitoring and accounting of both resources and
execution; and collection of static and sometimes dynamic information about
resources.

AliEn, DIRAC, GlideinWMS and PanDA all share a similar design with two types of
components: the management ones facing the application layer and centralizing
the capabilities required to acquire tasks' descriptions and matching them to
resource capabilities; and resource components required to acquire information
about resource capabilities and the resources themselves. Architecturally, the
management components include one or more queue and a scheduler capable of
coordinating via push/pull protocols with the resource modules. All resource
components include middleware-specific APIs for submitting resource requests
and, in the case of the pilot-enabled WMS of the LHC projects, a pilot capable
of pulling tasks from the management modules and scheduling them on their
resources (i.e., multi-level scheduling).

AliEn, DIRAC, GlideinWMS and PanDA present several similarities also in their
implementations. All were initially implemented to use Grid resources, using one
or more components to the Condor software ecosystem~\cite{thain2005distributed}
and, as with GlideinWMS, contributing to its development. Accordingly, all LHC
WMS implemented Grid-like authentication and authorization systems and adopted a
computational model based on distributing a large amount of single/few-cores
tasks across hundreds of sites\mtnote{Is this true?}. Tasks are assumed to fail
or being held and all WMS implements subsystems dedicated to manage these
failures or simply resubmitting failed tasks.

All the experiments at LHC produces and process large amount of data both from
actual collisions in the accelerator and from their simulations. Dedicated,
multi-tiered data systems have been built to store, replicate, and distributed
these data. All LHC WMS interface with these systems to move data to the sites
where related compute tasks are executed or to schedule compute tasks where
(large amount of) data are already stored.

Currently, the implementation of PanDA WMS has at least two distinguishing
features: the scale at which it manages computation, and network-aware
scheduling of tasks to sites. PanDA concurrently supports distributed
computations on up to 240,000 cores, submitting more than 55 million jobs in
2016 across more than 100 sites, and processing up to XXXXPB of data
distributed worldwide across XX Tier 1-3 facilities. As for 2017, PanDA serves
XXXX users, including XX production groups. The scale of this operation is
unprecedented, making PanDA the WMS managing the largest computational campaign
in the world.

PanDA was initially implemented to support computations across an unreliable and
relatively expensive networking infrastructure. Accordingly, PanDA assumed a
rigid model of data placement based on replication and long-term storage of data
on the Grid sites. This model proven too wasteful, requiring storage space for
unused and obsolete data. PanDA implementation was evolved to benefit from the
development of the internet and of the networking infrastructure for scientific
research. Today, PanDA constantly monitors network throughput and latency  among
its management components, Grid sites and data facilities. PanDA uses this
information to schedule tasks on sites, depending on the estimated time required
to download and/or replicate input data and to stage out output data.

The dynamic management of data replication and download is an example of a
generalized paradigm shift that distinguishes the implementation of PanDA from
other WMS. New components are being implemented to support the dynamic sizing of
input dataset based on resource capabilities, the dynamic sizing of the
resources held by pilots both in terms of number of cores and duration of the
pilot, and further abstraction of different types of resources to reconcile
High-Throughput and High-Performance Computing paradigms.

% Differences among workflow supported.

% AliEn, DIRAC, and PanDA are implemented to support specific experimental
% workflows, while GlideinWMS was developed as component of the Condor software
% framework. ALICE workload is \ldots. LHCb workload is \ldots. CMS workload is
% \ldots.
%
% The ATLAS project requires data, compute, and memory-intensive workflows, both
% for single users and so called `production' research groups. Generally, PanDA
% was designed to support mostly single-core tasks, requiring a variable \ldots.
%
% In this section, we present other workload manager systems that have been
% developed to support LHC related but they have been used also to support other
% computing demanding projects.

%\begin{itemize}
%   \item \emph{Clients}: consist in a set of APIs that allows users to submit
% job requests. Clients interact directly with DIRAC central services.
%   \item \emph{Resources}: they can be PC's, site clusters and Grids. Agents
% interact with them without distinction.
%\end{itemize}

% -----------------------------------------------------------------------------
\subsubsection{Alien}

Alien is a Workload and Data Management system composed of a set of middleware
tools and services entirely based on web-services and standard protocols. The
framework was originally developed for the ALICE experiment~\cite{Alice1995} but
subsequently used by several virtual organizations
\cite{McClatechey2003,GPCALMA}.

% The system has been deployed in 2001 for distributed production of Monte Carlo
% data, detector simulation and reconstruction.

Alien is composed of two type of services~\cite{Bagnasco2010}:

\begin{itemize}
    \item \emph{Central services}, these services are unique for each virtual organization, therefore there is only one configuration point for the management;
    \item \emph{Site services}, they provide the interfacing to local resources and Grid services running on a VO-box;
\end{itemize}

The most important Central services are:
\begin{itemize}
    \item \emph{Task Queue}, a database that keeps track of all the tasks submitted to the system and their current execution status;
    \item \emph{Brokers}, they are the core of task executions and data transfers; they receive tasks in form of JDL,  keep them ordered by priority and send them to the CE for execution;
    \item \emph{Optimizers}, they are used to minimize the work of the Broker by scanning periodially the task queue and re-arranging the tasks in such a way that fairness and priority policies are guaranted;
    \item \emph{Data Catalogue}, it keeps track of the scripts and files uploaded on Storage Elements.
\end{itemize}

%The Computing Agents are instead site services that monitor the local
% Computing Element, advertise site's capabilities and are responsible for
% submitting the JobAgents.

% Information about the status of the sites and central services, full job
% statistics and monitoring information are kept in a MonALISA repository.
%% CLUSTER MONITOR SHOULD BE EQUAL TO COMPUTING AGENT
%% Job Manager should be equal to TaskQueue
%% Process Monitor == PIlot????

The task execution in Alien is usually distributed over several sites. Each of
these sites has at least one service called ClusterMonitor. On one side the
Cluster Monitor is used to communicate with Central services (Task queue and
Broker), on the other side it can manage Computing Elements (CE) by starting and
stopping them whenever it receives the signal.

The CE is in charge of the execution of the tasks on the resources. A CE usually
is associated with a batch queue and can send the tasks to the worker nodes
controlled by the queue. The CE asks the Broker for tasks to execute by sending
its description in form of Job Description Language (JDL). Once received the
JDL, the Broker will try to match it with the JDL of the tasks in queue. If a
match exists then the Broker sends the tasks JDL to the CE. Immediately after
receiving a job JDL, the CE creates a new service on the worker node called
ProcessMonitor. This service allows the CE (and the rest of Alien services
through the CE) to interact with the job while is running \cite{Saiz2003}. This
execution strategy is called ``pull mode'' due to the fact that CE asks for
tasks.

% Since Alien exploits JDL, a workload can be described task by task according
% to features such as: task priorities, the level of parallelism (one core,
% multi-core, MPI etc..) and also the DCR that should be targeted for the
% execution.

% Job submission is implemented by following the so-called ``pull mode'' which
% is composed of the following steps:

% \begin{enumerate}
%   \item the VO-Box monitors the status of the site queues through polls to
% the resource running on the CE;
%   \item the Job Broker receives a report everytime slots become available;
%   \item if the Task Queue is not empty, the Job Broker asks the VO-box to
% submit a number of Agents;
%   \item finally, the JobAgents are submitted  to the site Computing Element
% either by way of that sends them back to the site Computing Element or,
% wherever available, directly through the CREAM interface on the CE itself.
% \end{enumerate}

% -----------------------------------------------------------------------------
\subsubsection{DIRAC}

DIRAC (Distributed Infrastructure with Remote Agent Control) Workload and Data
Management System has been developed within the CERN LHCb project to manage the
processing of detector data, Monte Carlo simulations, and end-user analyses
\cite{Tsaregorodtsev2004}. DIRAC's architecture relies on two entities
\cite{Paterson2010}:

\begin{itemize}
    \item \emph{Services}: serve Clients and Agents by performing crucial
    operations such as Task Management, Configuration, Bookkeeping and
    Accounting.
    \item \emph{Agents}: perform repetitive actions like querying file
    catalogs, monitoring of jobs on resources.
\end{itemize}

In the same way of Alien, DIRAC implements a pull scheduling. Furthermore DIRAC
was the first WMS to exploit the concept of Pilot Agent on the Grid
\cite{Casajus2010}, i.e., gLite jobs that are submitted to the grid when tasks
arrive into the WMS. DIRAC pilot system has four main logical components:

\begin{itemize}
    \item a set of TaskQueues that collect tasks submitted by users, multiple
    TaskQueue being created depending on the requirements and ownership of the
    tasks;
    \item a set of JobWrappers that are executed on the DCR to bind compute
    resources and execute tasks submitted by the users;
    \item a set of TaskQueueDirectors that submits JobWrappers to target DCRs;
    \item a MatchMaker that matches requests from JobWrappers to suitable tasks
    into TaskQueues.
\end{itemize}

The DIRAC execution model can be summarized in five steps: 1. a user submits its
workload in form of tasks to the WMS Job Manager; 2. submitted tasks are
validated and added to a new or an existing TaskQueue, depending on the task
properties; 3. TaskQueueDirector evaluates TaskQueues and a suitable number of
JobWrappers are submitted to available DCRs; 4. JobWrappers get instantiated on
the DCRs and, then, ask for tasks to the MatchMaker; 5. JobWrappers execute
tasks while JobWrapper’s Watchdog monitor them.

TaskQueueDirectors deploy Pilots by getting a list of TaskQueues and calculating
the number of pilot to submit according to user priorities. Once deployed on the
compute resource, Pilots, a.k.a. JobWrappers, hold the resource in the form of
single or multiple cores, spanning portions, whole, or multiple compute nodes.
Pilots do not expose data capabilities although the system allows the user to
perform both data staging and data replication. TaskQueues, TaskQueueDirectors,
and the MatchMaker are implemented as services whereas the JobWrapper is
implemented within the Agents together with the WatchDog.

% JDL and \emph{Transformation Management System} (TMS) are used for task
% description \cite{Tsaregorodtsev2006,Pacini2006}. The latter allows the user
% to describe complex data dependences.

% -----------------------------------------------------------------------------
\subsubsection{HTCondor Glidein and GlideinWMS}

The HTCondor Glidein system as part of the HTCondor software ecosystem. The
HTCondor Glidein is a pilot based system to aggregate DCRs with heterogeneous
middleware into HTCondor resource pools. Condor is based on daemons
collaborating by exchanging messages over the network. We can isolate four main
logical components~\cite{Sfiligoi2008}:

\begin{itemize}
    \item \emph{Schedd}, implements a queuing system that holds workload tasks;
    \item \emph{Startd}, controls the DCR resources.
    \item \emph{Collector}, holds references to all the activeSchedd/Startd daemons;
    \item \emph{Negotiator} matches tasks queued in a Schedd to resources handled by a Startd.
\end{itemize}

Glidein-WMS has been developed to integrate HTCondor Glidein to  automate the
deployment and management of Glideins on multiple types of DCR middleware. The
integration required three additional logical components:

\begin{itemize}
    \item \emph{Glidein Factories} that submit tasks to the DCRs middleware;
    \item a set of \emph{Virtual Organizations (VO) Frontend} daemons that
    match the tasks on one or more Schedd to the resource attributes;
    \item a \emph{Collector} that holds references to all the active Glidein
    Factories and VO Frontend daemons.
\end{itemize}

The execution model of the HTCondor Glidein system can be summarized in nine
steps: 1. the user submits a Glidein (i.e., a job) to a DCR batch scheduler; 2.
once executed, this Glidein bootstraps a Startd daemon; 3. the Startd daemon
advertises itself with the Collector; 4. the user submits the tasks of the
workload to the Schedd daemon; 5. the Schedd advertises these tasks to the
Collector; 6. the Negotiator matches the requirements of the tasks to the
properties of one of the available Startd daemon (i.e., a Glidein); 7. the
Negotiator communicates the match to the Schedd; 8. the Schedd submits the tasks
to the Startd daemon indicated by the Negotiator; 9. the task is executed.

By using GlideinWMS, the user does not have to submit Glidein directly but only
tasks to Schedd. From there: 1. every Schedd advertises its tasks with the VO
Frontend; 2. the VO Frontend matches the tasks’ requirements to the resource
properties advertised by the WMS Connector; 3. the VO Frontend places requests
for Glideins instantiation to the WMS Collector; 4. the WMS Collector contacts
the appropriate Glidein Factory to execute the requested Glideins; 5. the
requested Glideins become active on the DCRs; and 6. the Glideins advertise
their availability to the (HTCondor) Collector. From there on the execution
model is the same as described for the HTCondor Glidein Service.

% The resources managed by a single Glidein (i.e., pilot) are limited to compute
% resources. Glideins may bind one or more cores, depending on the target DCRs.
% For example, heterogeneous HTCondor pools with resources for desktops,
% workstations, small campus clusters, and some larger clusters will run mostly
% single core Glideins. More specialized pools that hold, for example, only DCRs
% with HTC, Grid, or Cloud middleware may instantiate Glideins with a larger
% number of cores. Both HTCondor Glidein and GlideinWMS provide abstractions for
% file staging but pilots are not used to hold data or network resources. The
% process of pilot deployment is the main difference between HTCondor Glidein
% and GlideinWMS. While the
%
% HTCondor Glidein system requires users to submit the pilots to the DCRs,
% GlideinWMS automates and optimizes pilot provisioning. GlideinWMS attempts to
% maximize the throughput of task execution by continuously instantiating
% Glideins until the queues of the available Schedd are emptied. Once all the
% tasks have been executed, the remaining Glideins are terminated. HTCondor
% Glidein and GlideWMS expose the interfaces of HTCondor to the application
% layer and no theoretical limitations are posed on the type and complexity of
% the workloads that can be executed. For example, DAGMan (Directed Acyclic
% Graph Manager) has been designed to execute workflows by submitting tasks to
% Schedd, and a tool is available to design applications based on the
% master-worker coordination pattern.
%
% Both HTCondor Glidein and GlideWMS rely on one or more HTCondor Collectors to
% match task requirements and resource properties, represented as ClassAds. This
% matching can be evaluated right before the scheduling of the task. In this
% way, late binding is achieved but early binding remains unsupported.
%
% \begin{table*}
% \begin{center}
% \begin{tabular}{llllllll}
%   \hline
%     Pilot System  &Logical Components& Execution Strategy & Binding  & Workload Definition  &  Broker  & \\
%   \hline
%     Alien & Central services, site services & pull, pilot via Co-pilot & Late & JDL & Condor (ClassAd attributes) &\\
%     DIRAC & Services, Agents& pull, pilot-based & Late & JDL, WF (TMS) & Condor(ClassAd attributes) &\\
%     glidein WMS& Daemons & pull, pilot-based  & Late & Pegaus, DAGMan & Condor(ClassAd attributes) &\\
%   \hline
% \end{tabular}
% \end{center}
% \caption{Comparison of the three Workload and Data Management Systems}\label{tab:Summary}
% \end{table*}