### 2017-03-26 appendix.tex
First iteration appendix template

Reproducibility initiative appendices: Artifact Description (AD) and Computational Results Analysis (CRA). Description of the organization of the appendix.

% -----------------------------------------------------------------------------
\subsection{Section 5.1 - Figure 3}
\label{apndx:fig3}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 5.1 - Figure 4}
\label{apndx:fig4}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 5.1 - Figure 5}
\label{apndx:fig5}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 5.2 - Figure 6}
\label{apndx:fig6}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 5.2 - Figure 7}
\label{apndx:fig7}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 5.3 - Figure 8}
\label{apndx:fig8}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 14}
\label{apndx:fig14}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 15}
\label{apndx:fig15}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 17}
\label{apndx:fig16}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 17}
\label{apndx:fig17}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}