### 2017-03-27 appendix.tex
Appendix Figure 8 and partial Figure 5

Reproducibility initiative appendices: Artifact Description (AD) and Computational Results Analysis (CRA). Description of the organization of the appendix.

% -----------------------------------------------------------------------------
\subsection{Section 5.1 - Figure 3}
\label{apndx:fig3}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

Replicability:

\begin{enumerate}
    \item
\end{enumerate}

% -----------------------------------------------------------------------------
\subsection{Section 5.1 - Figure 4}
\label{apndx:fig4}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

Replicability:

\begin{enumerate}
    \item
\end{enumerate}

% -----------------------------------------------------------------------------
\subsection{Section 5.1 - Figure 5}
\label{apndx:fig5}

Correlation between backfill availability of cores and walltime on Titan during the experiment time window.

\begin{itemize}
    \item Raw data: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.txt}, as provided by the PanDA Brokers logs on Titan.
    \item Wrangling: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.csv}
    \item Analysis and Plotting: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}
\end{itemize}

Replicability:

\begin{enumerate}
    \item
\end{enumerate}

% -----------------------------------------------------------------------------
\subsection{Section 5.2 - Figure 6}
\label{apndx:fig6}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

Replicability:

\begin{enumerate}
    \item
\end{enumerate}

% -----------------------------------------------------------------------------
\subsection{Section 5.2 - Figure 7}
\label{apndx:fig7}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

Replicability:

\begin{enumerate}
    \item
\end{enumerate}

% -----------------------------------------------------------------------------
\subsection{Section 5.3 - Figure 8}
\label{apndx:fig8}

Panda failures on Titan by exit code during the experiment window.

\begin{itemize}
    \item Raw data: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.csv}
    \item Wrangling: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Pivot Data \& Wrangled Data tabs)
    \item Analysis: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Aggregated Data tab)
    \item Plotting: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Plots tab)
\end{itemize}

Replicability:

\begin{enumerate}
    \item Download raw data download in csv format from the ATLAS jobs dashboard. Download link: \url{http://dashb-atlas-job.cern.ch/dashboard/request.py/terminatedjobsstatuscsv?sites=All%20T3210&sitesCat=All%20Countries&resourcetype=All&pandares=ORNL_Titan_MCORE&activities=simul&sitesSort=7&sitesCatSort=0&start=2016-01-01&end=2017-02-28&timeRange=daily&sortBy=16&granularity=Monthly&generic=0&series=30&type=abcb}
    \item Load the csv file into Excel (or any other software/language).
    \item Create a pivot table with failure types as columns and months as rows.
    \item Aggregate data of error codes of the same type by merge columns and adding the values of their raws according to this table:
    \item Plot the resulting table as a stacked area diagram.
\end{enumerate}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 14}
\label{apndx:fig14}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

Replicability:

\begin{enumerate}
    \item
\end{enumerate}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 15}
\label{apndx:fig15}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

Replicability:

\begin{enumerate}
    \item
\end{enumerate}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 17}
\label{apndx:fig16}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

Replicability:

\begin{enumerate}
    \item
\end{enumerate}

% -----------------------------------------------------------------------------
\subsection{Section 6.3 - Figure 17}
\label{apndx:fig17}

Artifacts description.

\begin{itemize}
    \item Experiment code: repository address (when applicable)
    \item Raw data: repository address
    \item Wrangling: repository address
    \item Analysis code: repository address
    \item Plotting code: repository address
\end{itemize}

Replicability:

\begin{enumerate}
    \item
\end{enumerate}