### 2017-04-04 PandaOverview.tex
Eliminate all comments and shift figure around to save space

PanDA is a Workload Management System (WMS) designed to support the execution of
distributed workloads and workflows via pilots. WMS is middleware for
discovering and selecting resources, submitting tasks of workloads and
workflows, and monitoring their execution~\cite{marco2009glite}. Pilot is an
abstraction that enables multi-level scheduling by decoupling resource
acquisition from tasks scheduling~\cite{turilli2015comprehensive}. When
implemented, a pilot is scheduled on a site and, once active, tasks are
scheduled to the pilot, not to the site's scheduler.

Pilot-capable WMS enable high throughput of tasks execution while supporting
interoperability across multiple sites. This is particularly relevant for LHC
experiments, where millions of tasks are executed across multiple sites every
month, analyzing and producing petabytes of data. The design of PanDA WMS
started in 2005 to support ATLAS. PanDA went into
production for the LHC Run 1 on 2009, and was then extended with new subsystems
for Run 2 in 2015.

% -----------------------------------------------------------------------------
\subsection{Design}
\label{ssec:panda_design}

PanDA's application model assumes tasks, workloads and workflows. Tasks
represent a set of operations performed on data sets stored in one or more input
files. Tasks are decomposed into jobs, where each job represents the task's set
of operations and a partition of the task's data set. Since 2005, a certain
amount of parallelism has been progressively introduced for job
execution~\cite{crooks2012multi} but, so far, HEP applications have not required
MPI. Jobs are supposed to be relatively self-contained, capable of setting up
their own execution environment or having a minimal set of common dependences.

PanDA's usage model is based on multitenancy of resources and the support of at
least two types of HEP users: individual researchers and groups executing so
called `production' workflows. Users are free to submit tasks and workflows to
the PanDA WMS at any point in time, directly or via dedicated application
frameworks. PanDA's security model is based on separation between
authentication, authorization and accounting for both single users and group of
users. Both authentication and authorization are based on digital certificates
and the X.509 standard and on the virtual organization (VO) abstraction.

Currently, PanDA's execution model is based on four main abstractions: task,
job, queue, and pilot. Both tasks and jobs are assumed to have attributes and
states and to be queued into a global queue for execution. Prioritization and
binding of jobs are assumed to depend on the attributes of each job. Pilot is
used to indicate the abstraction of resource capabilities. Each job is thought
to be bound to one pilot and executed on the site where the pilot has been
instantiated.

In PanDA's data model, each datum refers to the recorded or simulated
measurement of a physical process. Data can be packaged into files or other
containers. As with jobs, data have both attributes and states, and some of the
attributes are shared between events and jobs. Raw, reconstruction, and
simulation data are assumed to be distributed across multiple storage facilities
and managed by the ATLAS Distributed Data Management
(DDM)~\cite{garonne2012atlas}. When necessary, datasets required by each job are
assumed to be replicated over the network, both for input and output data.

PanDA's design supports provenance and traceability for both jobs and data.
Attributes enable provenance by linking jobs and data items, providing
information like ownership or project affiliation. States enable traceability by
providing information about the stage of the execution in which each job or data
item is or has been. Some attributes are assumed to be immutable across
execution and jobs and data items are assumed to be always in one and only one
state.

% -----------------------------------------------------------------------------
\subsection{Implementation and Execution}
\label{ssec:panda_arch}

The implementation of PanDA consists of several interconnected subsystems, most
of them built from off-the-shelf and Open Source components. Subsystems
communicate via dedicated API or HTTP messaging, and each subsystem is
implemented by one or more modules. Databases are used to store eventful
entities and entities' description like tasks, jobs and % events
input/output data, and to store information about sites,
resources, logs, and accounting.

Currently, PanDA's architecture has five main subsystems: PanDA
Server~\cite{maeno2011overview},
AutoPyFactory~\cite{caballero2012autopyfactory}, PanDA
Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of
ATLAS workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}), but we
do not discuss them as they are  not relevant to  an understanding of how
PanDA has been ported to supercomputers. For a full list of subsystems see
Ref.~\cite{panda-wiki_url}. Figure~\ref{fig:architecture} shows a diagrammatic
representation of PanDA main subsystems, highlighting the execution process of
tasks while omitting monitoring details to improve readability.

The relation between tasks and jobs can be one-to-one or one-to-many, and the
conversion between the two can by static or dynamic. During LHC Run 1, PanDA
required users or applications to perform a static conversion between tasks and
jobs: tasks were described as a set of jobs and then submitted to the PanDA
Server.

This approach introduced inefficiency both with usability and resource
utilization~\cite{borodin2015unified}. Ideally, users should not have to reason
in terms of jobs: Users conceive analyses in terms of one or more, possibly
related tasks; the `job' abstraction is required by the execution middleware,
i.e. PanDA. Further, a static partitioning of tasks into jobs does not take into
account the heterogeneity and dynamicity of the resources of the pilots on which
each job will be executed.

Another problem of static job sizing is that PanDA instantiates pilots on sites
with different type of resources and different models of availability of those
resources. An optimal sizing of each job should take into account these
properties. For example, sites may offer cores with different speed, networking
with different amount of bandwidth, and resources could be guaranteed to be
available for a defined amount of time or could disappear at any point in time
as with opportunistic models of resource provision.

JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users or
the application layer submit task descriptions to JEDI
(Fig.~\ref{fig:architecture}:1) that stores them into a queue implemented by a
database (Fig.~\ref{fig:architecture}:2). Tasks are partitioned into jobs of
different size, depending on both static and dynamic information about available
resources (Fig.~\ref{fig:architecture}:3). Jobs are bound to sites with
resources that best match jobs' requirements, and submitted to the PanDA Server
for execution (Fig.~\ref{fig:architecture}:4).

Once submitted to the PanDA Server, jobs are stored by the Task Buffer component
into a global queue implemented as a  database (Fig.~\ref{fig:architecture}:5).
When jobs are submitted directly to the PanDA Server, the Brokerage component is
used to bind jobs to available sites, depending on static information about the
resources available for each site. Jobs submitted by JEDI are already bound to
sites so no further brokerage is needed.

Once jobs are bound to sites, the Brokerage module communicates to the Data
Service module what data sets need to be made available on what site
(Fig.~\ref{fig:architecture}:6). The Data Service communicates these
requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7) that, when needed,
takes care of  replicating them on the required sites
(Fig.~\ref{fig:architecture}:8).

Meanwhile, AutoPyFactory defines a set of suitable PanDA Pilots and concurrently
submits them to a Condor-G broker (Fig.~\ref{fig:architecture}:9). This broker
submits these pilots wrapped as jobs or VMs to the required sites
(Fig.~\ref{fig:architecture}:10).

When a PanDA Pilot becomes available on a site, it calls the Job Dispatcher
module of the PanDA Server, requesting  a job to execute
(Fig.~\ref{fig:architecture}:11). The Job Dispatcher interrogates the Task
Buffer module for a job that is bound to the site of that pilot and ready to
be executed. Task Buffer checks the global queue (i.e., the PanDA database)
and, upon availability, returns a suitable job to the Job Dispatcher. The Job
Dispatcher dispatches that job to the PanDA Pilot
(Fig.~\ref{fig:architecture}:12).

Upon receiving a job, a PanDA Pilot starts a monitoring process and forks a
subprocess for the execution of the job's payload. The job is set up, input data
are transferred from the designated staging-in location
(Fig.~\ref{fig:architecture}:13), and the job's payload is executed
(Fig.~\ref{fig:architecture}:14). Once completed, output is transferred to the
staging-out location (Fig.~\ref{fig:architecture}:15).

The Data Service module of the PanDA Server tracks and collects the output
generated by each job (Fig.~\ref{fig:architecture}:16), updating jobs'
attributes via the Task Buffer module (Fig.~\ref{fig:architecture}:17). When the
output of all the jobs of a task are retrieved, it is made available to the user
via PanDA Server. When a task is submitted to JEDI, task is instead marked as
done (Fig.~\ref{fig:architecture}:18) and the result of its execution is made
available to the user by JEDI (Fig.~\ref{fig:architecture}:19).

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
    \caption{PanDA WMS architecture. Numbers indicates the execution process
    based on JEDI. Several subsystems and components, the architectural
    details of PanDA Pilot and the communication among subsystem's components
    are abstracted to improve clarity.}
\label{fig:architecture}
\end{figure}