### 2017-04-04 appendix.tex
Anonymize appendix

% -----------------------------------------------------------------------------
\subsection{Abstract}

\input{abstract}

% -----------------------------------------------------------------------------
\subsection{Description}

% -----------------------------------------------------------------------------
\subsubsection{Check-list (artifact meta information)}

{\small
\begin{itemize}
%  \item {\bf Algorithm: }
  \item {\bf Program: ATLAS Monte Carlo Workflow, Geant4, AthenaMP, GROMACS}
%  \item {\bf Compilation: }
%  \item {\bf Transformations: }
%  \item {\bf Binary: }
  \item {\bf Data set: Available at \\
  \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}}
  %\url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/tree/master/data}}
  \item {\bf Run-time environment: PanDA Workload Management System, NGE}
  % RADICAL-Pilot Pilot System (NGE)}
  \item {\bf Hardware: OLCF Titan Cray XK7}
%  \item {\bf Run-time state: }
%  \item {\bf Execution: }
  \item {\bf Output: Available at \\
  \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}}
  %\url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/tree/master/data}}
  \item {\bf Experiment workflow: Raw data acquisition, data wrangling and filtering, plotting, analysis}
%  \item {\bf Experiment customization: }
  \item {\bf Publicly available?: Yes}
\end{itemize}
}

% -----------------------------------------------------------------------------
\subsubsection{How software can be obtained (if available)}

\begin{itemize}
    \item PanDA Workload Management System: \\
    \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    % \url{https://github.com/PanDAWMS}.
    \item NGE: \\
    \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    % \url{https://github.com/radical-cybertools}.
\end{itemize}

% -----------------------------------------------------------------------------
\subsubsection{Hardware dependencies}

Access and allocation on OLCF Titan Cray XK7, workstation with at least 8GB of RAM.

% -----------------------------------------------------------------------------
\subsubsection{Software dependencies}

Python, jupyter, pandas, matplotlib, gnuplot, excel.

% -----------------------------------------------------------------------------
\subsubsection{Datasets}

\begin{description}[align=left]
    \item [Section 4.1 -- Figure 3] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 4.2 -- Figure 4] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 4.2 -- Figure 5] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 7] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 8] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 9] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
\end{description}

% -----------------------------------------------------------------------------
\subsection{Installation}

\begin{description}[align=left]
    \item [Section 4.1 -- Figure 3] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 4.2 -- Figure 4] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 4.2 -- Figure 5] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 7] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 8] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 9] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
\end{description}

% -----------------------------------------------------------------------------
\subsection{Experiment workflow}

\begin{description}[align=left]
    \item [Section 4.1 -- Figure 3] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 4.2 -- Figure 4] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 4.2 -- Figure 5] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 7] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 8] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 9] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
\end{description}

% -----------------------------------------------------------------------------
\subsection{Evaluation and expected result}

\begin{description}[align=left]
    \item [Section 4.1 -- Figure 3] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 4.2 -- Figure 4] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 4.2 -- Figure 5] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 7] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 8] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
    \item [Section 5.2 -- Figure 9] \url{URL_REMOVED_FOR_DOUBLE-BLIND_REVIEW}
\end{description}

% ----------------------------------------------------------------------------- % \subsection{Experiment customization}

% -----------------------------------------------------------------------------
% \subsection{Notes}

% =============================================================================
% \section{OLD -- REMOVE AFTER COMPLETING ABOVE}
% Reproducibility initiative appendices: Artifact Description (AD) and Computational Results Analysis (CRA). Description of the organization of the appendix.
%
% % -----------------------------------------------------------------------------
% \subsection{Section 5.1 - Figure 3}
% \label{apndx:fig3}
%
% Artifacts description.
%
% \begin{itemize}
%     \item Experiment code: repository address (when applicable)
%     \item Raw data: repository address
%     \item Wrangling: repository address
%     \item Analysis code: repository address
%     \item Plotting code: repository address
% \end{itemize}
%
% Replicability:
%
% \begin{enumerate}
%     \item
% \end{enumerate}
%
%
% % -----------------------------------------------------------------------------
% \subsection{Section 5.1 - Figure 4}
% \label{apndx:fig4}
%
% Artifacts description.
%
% \begin{itemize}
%     \item Experiment code: repository address (when applicable)
%     \item Raw data: repository address
%     \item Wrangling: repository address
%     \item Analysis code: repository address
%     \item Plotting code: repository address
% \end{itemize}
%
% Replicability:
%
% \begin{enumerate}
%     \item
% \end{enumerate}
%
%
% % -----------------------------------------------------------------------------
% \subsection{Section 5.1 - Figure 5}
% \label{apndx:fig5}
%
% Correlation between backfill availability of cores and walltime on Titan during the experiment time window.
%
% \begin{itemize}
%     \item Raw data: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.txt}, as provided by the PanDA Brokers logs on Titan.
%     \item Wrangling: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.csv}
%     \item Analysis and Plotting: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}
% \end{itemize}
%
% Replicability:
%
% \begin{enumerate}
%     \item Get a log file including regular polling of backfill availability from one or more PanDA Brokers on titan.
%     \item Wrangle the ascii file with the find/replace regular expressions listed at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}, cell \#1.
%     \item Load the Jupyter workbook from a python virtual environment with the module Jupyter installed and all the modules listed at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_5/titan_backfill_availability.ipynb}, cell \#2.
%     \item Execute each cell of the Jupyter notebook to replicate Figure 5.
% \end{enumerate}
%
%
% % -----------------------------------------------------------------------------
% \subsection{Section 5.2 - Figure 6}
% \label{apndx:fig6}
%
% Artifacts description.
%
% \begin{itemize}
%     \item Experiment code: repository address (when applicable)
%     \item Raw data: repository address
%     \item Wrangling: repository address
%     \item Analysis code: repository address
%     \item Plotting code: repository address
% \end{itemize}
%
% Replicability:
%
% \begin{enumerate}
%     \item
% \end{enumerate}
%
%
% % -----------------------------------------------------------------------------
% \subsection{Section 5.2 - Figure 7}
% \label{apndx:fig7}
%
% Artifacts description.
%
% \begin{itemize}
%     \item Experiment code: repository address (when applicable)
%     \item Raw data: repository address
%     \item Wrangling: repository address
%     \item Analysis code: repository address
%     \item Plotting code: repository address
% \end{itemize}
%
% Replicability:
%
% \begin{enumerate}
%     \item
% \end{enumerate}
%
%
% % -----------------------------------------------------------------------------
% \subsection{Section 5.3 - Figure 8}
% \label{apndx:fig8}
%
% Panda failures on Titan by exit code during the experiment window.
%
% \begin{itemize}
%     \item Raw data: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.csv}
%     \item Wrangling: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Pivot Data \& Wrangled Data tabs)
%     \item Analysis: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Aggregated Data tab)
%     \item Plotting: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/figure_8/panda-broker-failures_jan2016-feb2017.xlsx} (Plots tab)
% \end{itemize}
%
% Replicability:
%
% \begin{enumerate}
%     \item Download raw data download in csv format from the ATLAS jobs dashboard. Download link: \url{http://dashb-atlas-job.cern.ch/dashboard/request.py/terminatedjobsstatuscsv?sites=All%20T3210&sitesCat=All%20Countries&resourcetype=All&pandares=ORNL_Titan_MCORE&activities=simul&sitesSort=7&sitesCatSort=0&start=2016-01-01&end=2017-02-28&timeRange=daily&sortBy=16&granularity=Monthly&generic=0&series=30&type=abcb}
%     \item Load the csv file into Excel (or any other software/language).
%     \item Create a pivot table with failure types as columns and months as rows.
%     \item Aggregate data of error codes of the same type by merge columns and adding the values of their raws according to this table:
%     \item Plot the resulting table as a stacked area diagram.
% \end{enumerate}
%
%
% % -----------------------------------------------------------------------------
% \subsection{Section \ref{sec:ngeExp} - Figure \ref{fig:weakScal1a}}
% \label{apndx:fig9}
%
% \begin{itemize}
%     \item Raw data: \url{https://github.com/ATLAS-Titan/misc/tree/master/atlas-paper-experiments/data}
%      \item Processed data: \url{ https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/data.csv}
%     \item Raw data processing:
%     	\begin{itemize}
%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/dataAnalysis.py}
%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/script.py}
%     		\item Follow instructions at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-extract-data-from-the-sandbox}
%    	\end{itemize}
%     \item Plotting code: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/plot.gnu}
% \end{itemize}
%
% Replicability: Follow instructions at the following link: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-reproduce-experiments-on-TITAN} to prepare the environment on Titan.
% Launch:
% \begin{itemize}
% \item python dispatcher.py 250 250 120 etc/input\_100 to replicate experiments with 250 nodes;
% \item python dispatcher.py 500 500 120 etc/input\_100 to replicate experiments with 500 nodes;
% \item python dispatcher.py 1000 1000 120 etc/input\_100 to replicate experiments with 1000 nodes;
% \item python dispatcher.py 2000 2000 120 etc/input\_100 to replicate experiments with 2000 nodes.
% \end{itemize}
%
%
%
% % -----------------------------------------------------------------------------
% \subsection{Section \ref{sec:ngeExp} - Figure \ref{fig:weakScal2a}}
% \label{apndx:fig10}
% \begin{itemize}
%     \item Raw data: \url{https://github.com/ATLAS-Titan/misc/tree/master/atlas-paper-experiments/weak2}
%      \item Processed data: \url{ https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/weak2.csv}
%     \item Raw data processing:
%     	\begin{itemize}
%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/dataAnalysis.py}
%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/script.py}
%     		\item Follow instructions at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-extract-data-from-the-sandbox}
%    	\end{itemize}
%     \item Plotting code: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/plot.gnu}
% \end{itemize}
%
% Replicability: Follow instructions at the following link: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-reproduce-experiments-on-TITAN} to prepare the environment on Titan.
% Launch:
% \begin{itemize}
% \item python dispatcher.py 256 1280 256 180 etc/input\_16 to replicate experiments with 256 nodes;
% \item python dispatcher.py 512 2560 512 180 etc/input\_16 to replicate experiments with 512 nodes;
% \item python dispatcher.py 1024 5120 180 etc/input\_16 to replicate experiments with 1024 nodes;
% \item python dispatcher.py 2048 10240 180 etc/input\_16 to replicate experiments with 2048 nodes.
% \end{itemize}
%
%
% % -----------------------------------------------------------------------------
% \subsection{Section \ref{sec:ngeExp} - Figure \ref{fig:strongScala}}
% \label{apndx:fig16}
%
% \label{apndx:fig10}
% \begin{itemize}
%     \item Raw data: \url{https://github.com/ATLAS-Titan/misc/tree/master/atlas-paper-experiments/strong}
%      \item Processed data: \url{ https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/strong.csv}
%     \item Raw data processing:
%     	\begin{itemize}
%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/dataAnalysis.py}
%     		\item Download \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/script.py}
%     		\item Follow instructions at: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-extract-data-from-the-sandbox}
%    	\end{itemize}
%     \item Plotting code: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/blob/master/data/NGE/plot.gnu}
% \end{itemize}
%
% Replicability: Follow instructions at the following link: \url{https://github.com/ATLAS-Titan/PanDA-WMS-paper/wiki/How-to-reproduce-experiments-on-TITAN} to prepare the environment on Titan.
% Launch:
% \begin{itemize}
% \item python dispatcher.py 256  2048 256 180 etc/input\_16 to replicate experiments with 256 nodes;
% \item python dispatcher.py 512 2048 512 180 etc/input\_16 to replicate experiments with 512 nodes;
% \item python dispatcher.py 1024 2048 180 etc/input\_16 to replicate experiments with 1024 nodes;
% \item python dispatcher.py 2048 2048 180 etc/input\_16 to replicate experiments with 2048 nodes.
% \end{itemize}
%
%
%
% % -----------------------------------------------------------------------------
%