### 2017-04-04 PandaOverview.tex
Space-saving section 2

% PanDA is a Workload Management System (WMS) designed to support the execution of
% distributed workloads and workflows via pilots. WMS is middleware for
% discovering and selecting resources, submitting tasks of workloads and
% workflows, and monitoring their execution~\cite{marco2009glite}. Pilot is an
% abstraction that enables multi-level scheduling by decoupling resource
% acquisition from tasks scheduling~\cite{turilli2015comprehensive}. When
% implemented, a pilot is scheduled on a site and, once active, tasks are
% scheduled to the pilot, not to the site's scheduler.
%
% Pilot-capable WMS enable high throughput of tasks execution while supporting
% interoperability across multiple sites. This is particularly relevant for LHC
% experiments, where millions of tasks are executed across multiple sites every
% month, analyzing and producing petabytes of data. The design of PanDA WMS
% started in 2005 to support ATLAS. PanDA went into
% production for the LHC Run 1 on 2009, and was then extended with new subsystems
% for Run 2 in 2015.

PanDA is a Workload Management System (WMS)~\cite{marco2009glite} designed to
support the execution of distributed workloads and workflows via
pilots~\cite{turilli2015comprehensive}. Pilot-capable WMS enable high throughput
of tasks execution via multi-level scheduling while supporting interoperability
across multiple sites. This is particularly relevant for LHC experiments, where
millions of tasks are executed across multiple sites every month, analyzing and
producing petabytes of data. The design of PanDA WMS started in 2005 to support
ATLAS. PanDA went into production for the LHC Run 1 on 2009, and was then
extended with new subsystems for Run 2 in 2015.

% -----------------------------------------------------------------------------
\subsection{Design}
\label{ssec:panda_design}

PanDA's application model assumes tasks grouped into workloads and workflows.
Tasks represent a set of operations performed on data sets stored in one or more
input files. Tasks are decomposed into jobs, where each job consists of the
task's operations performed on a partition of the task's data set. Since 2005, a
certain amount of parallelism has been progressively introduced for job
execution~\cite{crooks2012multi} but, so far, HEP applications have not required
MPI.
% Jobs are supposed to be capable of setting up their own execution
% environment and have a minimal set of common dependences.

PanDA's usage model assumes multitenancy of resources and at least two types of
HEP users: individual researchers and groups executing so called `production'
workflows.
% Users can submit tasks and workflows to the PanDA WMS at any point in
% time, directly or via dedicated application frameworks.
PanDA's security model is based on separation between authentication,
authorization and accounting for both single users and group of users. Both
authentication and authorization are based on digital certificates and on the
virtual organization abstraction~\cite{foster2001anatomy}.

Currently, PanDA's execution model is based on four main abstractions: task,
job, queue, and pilot. Both tasks and jobs are assumed to have attributes and
states and to be queued into a global queue for execution. Prioritization and
binding of jobs are assumed to depend on the attributes of each job. Pilot is
used to indicate the abstraction of resource capabilities. Each job is thought
to be bound to one pilot and executed on the site where the pilot has been
instantiated.

In PanDA's data model, each datum refers to the recorded or simulated
measurement of a physical process. Data can be packaged into files or other
containers. As with jobs, data have both attributes and states, and some of the
attributes are shared between events and jobs. Raw, reconstruction, and
simulation data are assumed to be distributed across multiple storage facilities
and managed by the ATLAS Distributed Data Management
(DDM)~\cite{garonne2012atlas}. When necessary, datasets required by each job are
assumed to be replicated over the network, both for input and output data.

PanDA's design supports provenance and traceability for both jobs and data.
Attributes enable provenance by linking jobs and data items, providing
information like ownership or project affiliation. States enable traceability by
providing information about the stage of the execution in which each job or data
item is or has been.
% Some attributes are assumed to be immutable across execution and jobs and data
% items are assumed to be always in one and only one state.

% -----------------------------------------------------------------------------
\subsection{Implementation and Execution}
\label{ssec:panda_arch}

The implementation of PanDA WMS consists of several interconnected subsystems,
most of them built from off-the-shelf and Open Source components. Subsystems
communicate via dedicated API or HTTP messaging, and each subsystem is
implemented by one or more modules. Databases are used to store eventful
entities % and entities' description
like tasks, jobs and  input/output data, and to store information about sites,
resources, logs, and accounting.

Currently, PanDA's architecture has five main subsystems: PanDA
Server~\cite{maeno2011overview},
AutoPyFactory~\cite{caballero2012autopyfactory}, PanDA
Pilot~\cite{nilsson2011atlas}, JEDI~\cite{borodin2015scaling}, and PanDA
Monitoring~\cite{klimentov2011atlas}. Other subsystems are used by some of ATLAS
workflows (e.g., PanDA Event Service~\cite{calafiura2015atlas}), but we do not
discuss them as they are  not relevant to an understanding of how PanDA has been
ported to supercomputers. For a full list of subsystems see
Ref.~\cite{panda-wiki_url}. Fig.~\ref{fig:architecture} shows a diagrammatic
representation of PanDA main subsystems, highlighting the execution process of
tasks while omitting monitoring details to improve readability.

% The relation between tasks and jobs can be one-to-one or one-to-many, and the
% conversion between the two can by static or dynamic.

During LHC Run 1, PanDA required users to perform a static conversion between
tasks and jobs: tasks were described as a set of jobs and then submitted to the
PanDA Server. This introduced inefficiency both with usability and resource
utilization~\cite{borodin2015unified}. Ideally, users should conceive analyses
in terms of one or more, possibly related tasks, while the execution manager
(i.e., PanDA) should partition tasks into jobs, depending on execution
constraints. Further, the static partitioning of tasks into jobs does not take
into account the heterogeneity and dynamicity of the resources on which each job
will be executed, introducing inefficiencies.

Another problem of static job sizing is that PanDA instantiates pilots on sites
with different type of resources and different models of availability of those
resources. An optimal sizing of each job should take into account these
properties. For example, sites may offer cores with different speed, networking
with different amount of bandwidth, and resources could be guaranteed to be
available for a defined amount of time or could disappear at any point in time
as it happens with opportunistic models of resource provision.

JEDI was deployed for the LHC Run 2 to address these inefficiencies. Users
submit task descriptions to JEDI (Fig.~\ref{fig:architecture}:1) that stores
them into a queue implemented by a database (Fig.~\ref{fig:architecture}:2).
Tasks are partitioned into jobs of different size, depending on both static and
dynamic information about available resources (Fig.~\ref{fig:architecture}:3).
Jobs are bound to sites with resources that best match jobs' requirements, and
submitted to the PanDA Server for execution (Fig.~\ref{fig:architecture}:4).

Once submitted to the PanDA Server, jobs are stored by the Task Buffer component
into a global queue implemented as a database (Fig.~\ref{fig:architecture}:5).
When jobs are submitted directly to the PanDA Server, the Brokerage component is
used to bind jobs to available sites, depending on static information about the
resources available for each site. Jobs submitted by JEDI are already bound to
sites so no further brokerage is needed.

Once jobs are bound to sites, the Brokerage module communicates to the Data
Service module what data sets need to be made available on what site
(Fig.~\ref{fig:architecture}:6). The Data Service communicates these
requirements to the ATLAS DDM (Fig.~\ref{fig:architecture}:7) that, when needed,
replicates data sets on the required sites (Fig.~\ref{fig:architecture}:8).

Meanwhile, AutoPyFactory defines PanDA Pilots, submitting them to a Condor-G
agent (Fig.~\ref{fig:architecture}:9). Condor-G schedules these pilots wrapped
as jobs or VMs to the required sites (Fig.~\ref{fig:architecture}:10).

When a PanDA Pilot becomes available, it requests the Job Dispatcher module of
the PanDA Server for a job to execute (Fig.~\ref{fig:architecture}:11). The Job
Dispatcher interrogates the Task Buffer module for a job that is bound to the
site of that pilot and ready to be executed. Task Buffer checks the global queue
(i.e., the PanDA database) and, upon availability, returns a job to the Job
Dispatcher. The Job Dispatcher dispatches that job to the PanDA Pilot
(Fig.~\ref{fig:architecture}:12).

Upon receiving a job, a PanDA Pilot starts a monitoring process and forks a
subprocess for the execution of the job's payload. Input data are transferred
from the stage-in location (Fig.~\ref{fig:architecture}:13) and the job's
payload is executed (Fig.~\ref{fig:architecture}:14). Once completed, output is
transferred to the staging-out location (Fig.~\ref{fig:architecture}:15).

The Data Service module of the PanDA Server tracks and collects the output
generated by each job (Fig.~\ref{fig:architecture}:16), updating jobs'
attributes via the Task Buffer module (Fig.~\ref{fig:architecture}:17). When the
output of all the jobs of a task are retrieved, it is made available to the user
via PanDA Server. When a task is submitted to JEDI, task is instead marked as
done (Fig.~\ref{fig:architecture}:18) and the result of its execution is made
available to the user by JEDI (Fig.~\ref{fig:architecture}:19).

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_architecture.pdf}
    \caption{PanDA WMS architecture. Numbers indicates the JEDI-based execution
    process. Several subsystems, components, and architectural and
    communication details are abstracted for clarity.}
\label{fig:architecture}
\end{figure}