### 2017-04-04 overall.tex
Small iteration 4

Currently, ATLAS deploys 20 instances of the PanDA Broker on 4 Titan's DTNs, 5
instances per DTN. Each broker submits and manages the execution of 15 to 300
ATLAS jobs, one job for each Titan's worker node, and a theoretical maximum
concurrent use of 96,000 cores. Since November 2015, PanDA Brokers have operated
only in backfill mode, without a defined time allocation, and running at the
lowest priority on Titan.

We evaluate the efficiency, scalability and reliability of the deployment of
PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and
AthenaMP. All the measurements were performed between January 2016 and February
2017, hereafter called `experiment time window'.

% -----------------------------------------------------------------------------
\subsection{Characterizing the PanDA Broker on Titan}
\label{ssec:broker_titan}

We calculate the total amount of backfill availability over a period of time by:
(i) polling the available backfill slots at regular intervals during that time
window; (ii) converting the number of worker nodes available and their walltime
into core-hours; (iii) summing the number of core-hours. We call this number of
core-hours `total backfill availability'.

Fig.~\ref{fig:backfill-utilization} shows the total backfill availability on
Titan (gray bars) and the number of core-hours of that availability used by
ATLAS (blue bars) during the experiment time window. ATLAS consumed a total of
51.4M core-hours, for an average of $\approx$3.7M core-hours a month, with a
minimum of 1.7M core-hours in April 2016 and a maximum 7.9M core-hours in
February 2017.

\begin{figure}[!t]
    \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.pdf}
    \vspace{-0.3in}
    \caption{Titan's total backfill availability: CPU core-hours (gray) and CPU
    core-hours used by ATLAS (blue). GPU core-hours unaccounted for as they
    cannot be used by ATLAS. Efficiency of PanDA Brokers defined as percentage
    of total Titan's backfill availability used by ATLAS (Red labels).}
\label{fig:backfill-utilization}
\end{figure}

PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels) is
defined as the fraction of core-hours utilized by the PanDA Brokers to  Titan’s
total backfill availability during the experiment time window. The brokers
reached 18\% average efficiency, with a minimum 7.8\% efficiency on May 2016 and
a maximum 30.9\% efficiency on February 2017 (excluding the preliminary results
of March). The total backfill availability was 20.3M in April 2016, and 17.6M in
February 2017. This shows that the measured difference in efficiency did not
depend on a comparable difference in the total backfill availability.

During the experiment time window, about 2.25M detector simulation jobs were
completed on Titan, for a total of 225M events processed. This is equivalent to
0.9\% of all the 250M detector simulations performed by ATLAS in the same period
of time, and 3.5\% of the 6.6B events processed by those jobs. These figures
confirms the relevance of supercomputers' resource contribution to the LHC Run
2, especially when accounting for the amount of unused total backfill
availability and the improvement of PanDA efficiency across the experiment time
window.

On February 2017, PanDA Brokers used almost twice as much total backfill
availability than in any other month (preliminary results for March 2017
displayed in Fig.~\ref{fig:backfill-utilization} confirm this trend). No
relevant code update was made during that period and logs indicated that the
brokers were able to perform faster. This is likely due to hardware upgrades on
the DTNs. The absence of continuous monitoring of those nodes does not allow to
quantify bottlenecks but spot measurements of their load indicate that a faster
CPU and better networking were likely responsible for the improved performance.
Investigations showed an average CPU load of 3.6\% on the upgraded DTNs. As
such, further hardware upgrades seem unlikely to improve significantly the
performance of PanDA Brokers.

Every detector simulation executed on Titan process 100 events. This number of
events is consistent with the physics of the use case and with the average
duration of backfill availability. The duration of a detector simulation is a
function of the number of events simulated but not all events take the same time
to be simulated. 1 event simulation takes from $\approx$2 to $\approx$40
minutes, with a mean of $\approx$14 minutes. Considering that each worker node
process up to 16 events concurrently, 100 events takes an average of 105 minutes
to process. As such, PanDA brokers do not use backfill availability with less
than 105 minutes walltime.

Fig.~\ref{fig:backfill-distrib} shows the distribution of backfill availability
on Titan as a function of number of nodes and the time of their availability
(i.e., walltime). We recorded these data by polling Titan's Moab scheduler at
regular intervals during the experiment window time. The mean number of nodes
was 691, and their mean walltime was 126 minutes. Detector simulations of 100
events, enable to use down to 5/6 of the mean walltime of backfill availability.
As such, it offers a good compromise for PanDA Broker efficiency.

\begin{figure*}[!t]
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_wnodes_distribution.pdf}
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_walltime_distribution.pdf}
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_avail.png}
   \vspace{-0.1in}
    \caption{Backfill availability: distribution of number of work nodes (left)
    and of walltime in minutes (center). Scatter plot of the two variables
    (right). 62175 measures: mean number of work nodes available 691; mean
    walltime available 126 minutes.}

\label{fig:backfill-distrib}
\end{figure*}

% Usually, detector simulations performed on the Grid process 10 times the
% number of events than those on Titan. This explains the difference between the
% percentage of detector simulations and of events computed by those simulations
% on Titan.
PanDA Broker could fit the number of events to the walltime of each available
backfill slot on the base of the distributions of the time taken by one event to
be simulated. That specific number of event could then be pulled from the PanDA
Event service~\cite{calafiura2015atlas} and given as input to one or more
simulations. Once packaged into the MPI script submitted to titan's PBS batch
system, these simulations would better fit their available backfill slot,
contributing to increase the efficiency of PanDA Brokers.

The transition from a homogeneous to a heterogeneous number of events per
detector simulation has implications for the application layer. An even number
of events across simulations makes it easier to partition, track and package
events across simulations, especially when they are performed on both the Grid
and Titan. A homogeneous number of events also helps to keep the size and
duration of other stages of the MC workflow (\S\ref{ssec:panda-titan}) more
uniform. Further analysis is needed to evaluate the trade offs between increased
efficiency of resource utilization and the complexity that would be introduced
at the application layer.

Currently, each PanDA Broker creates, submits, and monitors a single MPI PBS
script at a time. This design is inherited from PanDA Pilot where a single
process is spawn at a time to execute the payload. As a consequence, the
utilization of a larger portion of Titan's total backfill availability depends
on the the number of concurrent PanDA Brokers instantiated on the DTNs: When all
the 20 PanDA Brokers have submitted a MPI PBS script, further backfill
availability cannot be used.

In September 2016, increasing the number of concurrent PanDA brokers from 4 to
20 markedly improved efficiency (see Fig.~\ref{fig:backfill-utilization}) but
further research is undergoing to understand whether an even larger number of
brokers would yield similar results. This research focuses on evaluating the
overheads of input/output files staging, including its impact on DTNs, and on an
alternative design of PanDA Broker that enables the concurrent submission of
multiple MPI scripts~\cite{barreiro2016panda}.

The current design and architecture of the PanDA Broker is proving to be as
reliable as PanDA Pilot when used on the WLCG. Between Jan 2016 and Feb 2017,
the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
while the failure rate of jobs submitted to Titan was a comparable 13.6\%. PanDA
Brokers were responsible for around the 19\% of the failures, compared to the
29\% of failures produced by the JobDispatcher module of the PanDA Server, and
the 13\% failures produced by the Geant4 toolkit. The current failure rate of
the PanDA Brokers confirms the benefits of reusing most of the code base of the
PanDA Pilot for implementing the PanDA Broker. It also shows that adopting
third-party libraries like RADICAL-SAGA did not have a measurable adverse effect
on reliability.

% -----------------------------------------------------------------------------
\subsection{Characterizing the Detector Simulation on Titan}
\label{ssec:athenamp_titan}

We use two main parameters to measure the performance of the detector simulation
jobs submitted to Titan: (i) the time taken to setup AthenaMP; and (ii) the
distribution of the time taken by the Geant4 toolkit to simulate a certain
number of events.

AthenaMP has an initialization and configuration stage. At initialization time,
AthenaMP is assembled from a large number of shared libraries, depending on the
type of payload that will have to be computed. Once initialized, every algorithm
and service of AthenaMP is configured by a set of Python scripts. Both these
operations result in a large number of read operations on the filesystem shared
between the worker nodes and the DTNs, including the operations required to
access small python scripts.

Initially, all the shared libraries of AthenaMP and the python scripts for its
configuration were stored on the Spider 2 Lustre file system. However, the I/O
patterns of the initialization and configuration stages degraded the performance
of the filesystem. This was addressed by moving the AthenaMP distribution to a
read-only NFS directory, shared among DTNs and worker nodes. NFS eliminated the
problem of metadata contention, improving metadata read performance from
$\approx$6,300 seconds on Lustre to $\approx$1,500 seconds on NFS.

Once initialized and configured, AthenaMP is used to execute 16 concurrent
Geant4 simulators on a Titan's worker node. Geant4 requires to read events
descriptions from a filesystem and simulate them as they would happen within the
ATLAS detector. We characterized both the compute performance of the simulation
and the impact of acquiring event descriptions on the filesystem.

The AMD Opteron 6274 CPU used on Titan has 16 cores, divided into 8 compute
units. Each compute units has 1 floating point (FP) scheduler shared between 2
cores. When using 16 cores for FP-intensive calculations, each pair of cores
competes for a single FP scheduler. This creates the overhead shown in
Fig.~\ref{fig:comparison-8-16cores}: the mean runtime per event for 8
concurrent simulations computing 50 events is 10.8 minutes, while for 16
simulations is 14.25 minutes (consistent with the measured distribution of the
duration of event simulation). Despite an inefficiency of almost 30\%, Titan's
allocation policy based on number of worker nodes used instead of number of
cores does not justify the use of 1/2 of the cores available.

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison_vsquashed.pdf}
   \vspace{-0.3in}
    \caption{Distributions of the time taken to simulate one event when placing
    2 simulations (h1) or 1 simulation (h2) per Titan's CPU. 2 simulation use 16
    cores per node, 1 simulation 8. 50 Events; 1 Titan worker nodes; 16 work
    threads per node; 100 events per node.}

\label{fig:comparison-8-16cores}
\end{figure}

The performance analysis of Titan's AMD CPUs for detector simulations helps also
to compare Titan and Grid site performances. Usually, Grid sites exposes
resources with heterogeneous CPU architectures and a maximum of 8 (virtual)
cores per worker node, while Titan's offer an homogeneous 16 cores architecture.
We used the rate of events processes per minute as a measure of the efficiency
of executing the same detector simulation on Titan or Grid sites. Comparisons of
the efficiencies of Titan to the BNL and SIGNET Grid sites, normalized for 8
cores, show that the effective performance per-core at Titan is $\approx$0.57
event per minute, roughly 1/2 of BNL and  1/3 of SIGNET performances.

The differences in performance between Titan and the BNL and SIGNET Grid sites
are due to the FP scheduler competition and the availability of newer
processors. The CPUs at the Grid sites have one FP scheduler per core and are on
average newer than the CPU of Titan. The heterogeneity of the Grid sites' CPUs
explain the higher performance variance we observed compared to the performance
consistency we measured on Titan.

We studied the impact of acquiring event descriptions on Lustre by analyzing
1,175 jobs ran on the week of 10/25/2016, for a total of 174 hours.
Table~\ref{panda-olcf-stats} shows the overall statistical breakdown of the
observed file I/O. ATLAS used between 1 and 300 worker nodes, and 35 on average.
75\% of the jobs run by ATLAS consumed less than 25 nodes and 92\% less than
100. During the 174 hours of data collection, 6.75 ATLAS jobs were executed on
average per hour, each job running for an average of 1.74 hours. Every job read
less than 250 GB and wrote less than 75 GB of data and, on average, each job
read 20 GB and wrote 6 GB of data.

\begin{table*}[t]
\centering
\begin{tabular}{lllllllll}
 & Num. Nodes & Duration (s) & Read (GB) & Written (GB) & GB Read/nodes & GB Written/nodes & $open()$ & $close()$ \\
Min & 1 & 1,932 & 0.01 & 0.03 & 0.00037 & 0.02485 & 1,368 & 349 \\
Max & 300 & 7,452 & 241.06 & 71.71 & 0.81670 & 0.23903 & 1,260,185 & 294,908 \\
Average & 35.66 & 6,280.82 & 20.36 & 6.87 & 0.38354 & 0.16794 & 146,459.37 & 34,155.74 \\
Std. Dev. & 55.33 & 520.99 & 43.90 & 12.33 & 0.19379 & 0.03376 & 231,346.55 & 53,799.08
\end{tabular}
\caption{The Statistical breakdown of the I/O impact of 1,175 jobs ATLAS executed at OLCF for the week of 10/25/16}
\label{panda-olcf-stats}
\end{table*}

ATLAS jobs are read heavy: On average, the amount of data read per worker node
is less than 400 MB, while the amount of data written is less than 170 MB.
Distributions of read and written data are different: The read operation
distribution per job shows a long tail, ranging from 12.5 GB to 250 GB, while
the written amount of data has a very narrow distribution.

The metadata I/O breakdown shows that ATLAS jobs yield 23 file $open()$
operations per second (not including file $stat()$ operations) and 5 file
$close()$ operations per second, with similar distributions. On average, the
maximum number of file $open()$ operations per job is $\approx$170/s and the
maximum number of file $close()$ operations is $\approx$39/s. For the 1,175
ATLAS jobs observed, the total number of file $open()$ operations is 172,089,760
and the total number of file $close()$ operations is 40,132,992. The difference
between these two values is still under investigation: One possible explanation
is that ATLAS jobs don't call a file $close()$ operation per every file $open()$
issued.

Overall, the average time taken to read events from input files stored on Lustre
is 1,320, comparable to the time taken to read the file required by assembling
AthenaMP from NFS. Preliminary investigation shows that this time could be
reduced to 40 seconds by loading the event descriptions into the RAM disk
available on each worker node. Events descriptions could be transferred from
Lustre to the RAM disk while configuring and initializing AthenaMP, almost
halving the time currently required by initiating a Genat4 simulation.