### 2017-04-04 conclusion.tex
conclusions. refinements.

% PanDA WMS was developed to meet the scale and complexity of LHC
% distributed computing for the ATLAS experiment. In the process, the old batch
% job paradigm of computing in HEP was discarded in favor of a far more
% flexible and scalable model. The success of PanDA at the LHC is leading to
% widespread adoption and testing by other experiments. PanDA is the first
% exascale workload management system in HEP, already operating at a million
% computing jobs per day, and processing over an exabyte of data in 2013. Next
% LHC run will pose massive computing challenges. With a doubling of the beam
% energy and luminosity as well as an increased need for simulates data, the
% data volume is expected to increase with a factor 5--6 or more. Storing and
% processing this amount of data is a challenge that cannot be resolved with the
% currently existing computing resources in ATLAS\@. To resolve this challenge,
% ATLAS is turning to commercial as well as academic Cloud services and HPCs via
% the PanDA system. Also the work underway is enabling the use of PanDA by new
% scientific collaborations and communities as a means of leveraging extreme
% scale computing resources with a low barrier of entry. The technology base
% provided by the PanDA system will enhance the usage of a variety of
% high-performance computing resources available to basic research.

The deployment of PanDA Broker on Titan enabled distributed computing on a
leadership-class HPC machine at unprecedented scale. In the past 13 months,
PanDA WMS has consumed almost 52M core-hours on Titan, simulating 3.5\% of the
total number of detector events of the ATLAS production Monte Carlo workflow. We
described the implementation and execution process of PanDA WMS
(\S\ref{sec:panda_overview}) and PanDA Broker (\S\ref{sec:panda_titan}), showing
how they support and enable distributed computing at this scale on Titan, a
leadership-class HPC machine managed by OCLF at ORNL.

We characterized the state of practice by evaluating the efficiency, scalability
and reliability of both PanDA Broker and AthenaMP as deployed on Titan
(\S\ref{sec:panda_titan}). Our characterization highlighted the strengths
and limitations of the current design and implementation: PanDA Brokers enable
the sustained execution of millions of simulations per week but further work is
required to optimize its efficiency and reliability (\S\ref{ssec:broker_titan}).
PanDA Brokers support the concurrent execution of multiple AthenaMP instances,
enabling each AthenaMP to perform the concurrent execution of up to 16 Geant4
simulators. Nonetheless, our characterization showed how improving I/O
performance could reduce overheads (\S\ref{ssec:athenamp_titan}), increasing the
overall utilization of Titan's backfill availability.

More generally, this paper highlights the fundamental role of WMS for
experimental science. HEP was amongst the first, if not the very first
experimental community to realize the importance of using WMS to manage their
computational campaign(s). As computing becomes increasingly critical for a
range of experiments, the impact of WMS on HEP foreshadows the importance of
WMS for other experiments (such as SKA, LSST etc.). % and their computational
campaigns. These experiments will have their own workload characteristics,
resources types and federation constraints, as well metrics of performance.
The design and experience captured in this paper will prove invaluable for
designing WMS to support computational campaigns and will provide a baseline
to evaluate the relative merits of different approaches.

The 52M core hours used by ATLAS, via PanDA, is over 2\% of the total
utilization on Titan over the same period, bringing the time-averaged
utilization of Titan to be consistently upwards of 90\%. Given that the
average average utilization of most other leadership class machines is less
(e.g., NSF's flagship Blue Waters the average utilization fluctuates between
60-80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
approaches elsewhere. These unprecedented efficiency gains aside, this work is
just a starting point towards more effective operational models for future
leadership and online analytical platforms~\cite{foap-url}. These platforms
will have to support ever increasing complex workloads with varying models for
dynamic resource federation. Recent efforts with Google Cloud for the CMS
experiment -- HEPCloud~\cite{hepcloud,googlehep} represent similar yet
different approaches to dynamic resource federation. As the type,
heterogeneity and complexity of future platforms  (e.g., Exascale, Commercial
Clouds) for science increases, there will be a greater need and emphasis on
abstractions to ensure WMS can both utilize as well as support evolving
platforms. Our experience demonstrated the advantages arising from an
abstractions based approach \textendash{} in particular scalable implementations
of pilot abstraction.