### 2017-06-26 nge.tex
minor reduction.

As seen in \S\ref{ssec:panda_titan}, PanDA Broker was  deployed on the DTNs
as Titan's worker nodes lack connectivity to the wide area network The lack
of pilot capabilities impacts both the efficiency and the flexibility of
PanDA's execution process. Pilots could improve efficiency by increasing
throughput and enabling greater backfill utilization. Further, pilots makes
it easier to support heterogeneous workloads.

The absence of pilots imposes the static coupling between MPI scripts
submitted to the PBS batch system and detector simulations
(\S\ref{sec:panda_titan}). This makes the scheduling of multiple generations
of workload on the same PBS job impossible: once a statically defined number
of detector simulations are packaged into a PBS job and this job is queued on
Titan, no further simulations can be added to that job. New simulations have
to be packaged into a new PBS job that needs to be submitted to Titan based
upon backfill availability.

The support of  multiple generations of workload would enable more efficient
use of the backfill availability of walltime. Currently, when a set of
simulations ends, the PBS job also ends, independent of whether more
wall-time would still be available. With a pilot, additional simulations
could be executed  to utilize all the available wall-time, while avoiding
further job packaging and submission overheads.

Multiple generations would also relax two assumptions of the current
execution model: knowing the number of simulations before submitting the MPI
script, and having a fixed number of events per simulation (100 at the
moment). Pilots would enable the scheduling of simulations independently from
whether they were available at the moment of submitting the pilot. Further,
simulations with a varying number of events could be scheduled on a pilot,
depending on the amount of remaining walltime and the distribution of
execution time per event, as shown in \S\ref{ssec:panda_titan},
Fig.~\ref{fig:comparison-8-16cores}. These capabilities would increase the
efficiency of the PanDA Broker when there is a large difference between the
number of cores and walltime.

Pilots can offer a payload-independent scheduling interface while hiding the
mechanics of coordination and communication among multiple worker nodes. This
could eliminate the need for packaging payload into MPI scripts within the
broker, greatly simplifying the submission process. This simplification would
also enable the submission of different types of payload, without having to
develop a specific PBS script for each payload. The submission process would
also be MPI-independent, as MPI is used for coordination among multiple
worker nodes, not by the payload.

% -----------------------------------------------------------------------------
\subsection{Implementation}
\label{sec:arch}

The implementation of pilot capabilities within the current PanDA Broker require
quantification of the effective benefits that it could yield and, on the base of
this analysis, a dedicated engineering effort. We developed a prototype of a
pilot system capable of executing on Titan to study experimentally the
quantitative and qualitative benefits that it could bring to PanDA. We called
this prototype Next Generation Executor (NGE).

NGE is a runtime system to execute heterogeneous and dynamically determined
tasks that constitute workloads. Fig.~\ref{fig:arch-overview} illustrates its
current architecture as deployed on Titan: the two management modules (Pilot
and Unit) represent a simplified version of the PanDA Broker while the agent
module is the pilot submitted to Titan and executed on its worker nodes. The
communication between PanDA Broker and Server is abstracted away as it
is not immediately useful to evaluate the performance and capabilities of a
pilot on Titan.

\begin{figure}
  \centering
   \includegraphics[width=\columnwidth]{figures/rp_architecture_compact_atlaswms_paper.pdf}
   \vspace{-0.3in}
  \caption{NGE Architecture: The PilotManager and  UnitManager reside on a Titan
  DTN while the Agent is executed on its worker nodes. Color coding: gray for
  entities external to NGE, white for APIs, purple for NGE's modules, green for
  pilots, yellow for module's components.}
\label{fig:arch-overview}
\end{figure}

NGE exposes an API to describe workloads (Fig.~\ref{fig:arch-overview}, green
squares) and pilots (Fig.~\ref{fig:arch-overview}, red circles), and to
instantiate a PilotManager and a UnitManager.   The PilotManager submits
pilots to Titan's PBS  batch system via SAGA API (Fig.~\ref{fig:arch-
overview}, dash arrow), as done by PanDA Broker to submit MPI scripts. Once
scheduled, the pilot's Agent is bootstrapped on Titan's MOM node and the
Agent's Executors on the on the worker nodes, and the UnitManager schedules
units to the Agent's Scheduler (Fig.~\ref{fig:arch-overview}, solid arrow).
The Agent's Scheduler schedules the units on one or more Agent's Executor for
execution. The Agent's executors can manage one or more worker nodes,
depending on performance evaluations. The UnitManager and the Agent
communicate via a database that is instantiated on one of Titan's DTN so as to
be reachable by both modules.

The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for
communication and coordination of the execution of units. This environment is
a critical component of the OpenMPI implementation ~\cite{castain05:_open_rte,
cug-2016}. ORTE is able to minimize the system overhead while submitting tasks
by avoiding filesystem bottlenecks and race conditions with network sockets.

% The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for
% communication and coordination of the execution of units. This environment is
% a critical component of the OpenMPI implementation, developed to support
% distributed high-performance computing applications operating in a
% heterogeneous environment~\cite{castain05:_open_rte, cug-2016}. ORTE provides
% a mechanism to create a \emph{``dynamic virtual machine''} (DVM) that spans
% multiple nodes. The DVM transparently provides support for interprocess
% communication, resource discovery and allocation, and process launching across
% a variety of platforms. Libraries enable the interaction between users and
% ORTE to enable the submission, monitoring and managing of tasks, avoiding
% filesystem bottlenecks and race conditions with network sockets. As a
% consequence, ORTE is able to minimize the system overhead while submitting
% tasks.