### 2017-06-27 paper.tex

\documentclass[conference]{IEEEtran}

% ---------------------------------------------------------------------------
% Authors are invited to submit unpublished, original work, using the IEEE
% 8.5 × 11 manuscript guidelines: double-column text using single-spaced 10
% point font on 8.5 × 11 inch pages. Templates are available from http://www.
% ieee.org/conferences_events/conferences/publishing/templates.html.
%
% The conference proceedings will be made available online through the IEEE
% Digital Library.
%
% Three types of papers will be accepted:
%
% - Full research papers (up to 10 pages): focus on new research achievements
%   in eScience, spanning core technologies and applications.
% - Experience papers (up to 10 pages): focus on practical outcomes through
%   applying eScience tools and techniques to novel problems.
% ---------------------------------------------------------------------------

\input{head}
\input{include}

% ---------------------------------------------------------------------------
% BEGIN DOCUMENT
% ---------------------------------------------------------------------------
\begin{document}

\title{Converging High-Throughput and High-Performance Computing: A Case Study}

\author{
    \IEEEauthorblockN{
        Alessio Angius\IEEEauthorrefmark{1}\IEEEauthorrefmark{2},
        Danila Oleynik\IEEEauthorrefmark{1}\IEEEauthorrefmark{5},
        Sergey Panitkin\IEEEauthorrefmark{1}\IEEEauthorrefmark{4},
        Matteo Turilli\IEEEauthorrefmark{1}\IEEEauthorrefmark{2},\\
        Kaushik De\IEEEauthorrefmark{5},
        Alexei Klimentov\IEEEauthorrefmark{4},
        Sarp H. Oral\IEEEauthorrefmark{6},
        Jack C. Wells\IEEEauthorrefmark{6} and
        Shantenu Jha\IEEEauthorrefmark{2}
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{1}First authors, alphabetical order
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{2}Department of Electric and Computer Engineering, Rutgers University, NJ, USA
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{4}Physics Department, Brookhaven National Laboratory, NY, USA
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{5}Physics Department, University of Texas Arlington, TX, USA
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{6}Oak Ridge Leadership Computing Facility, Oak Ridge National Laboratory, TN, USA
    }
}

% ---------------------------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------------------------
\maketitle
\begin{abstract}
\input{abstract}
\end{abstract}
% \keywords{Distributed computing, Workload management system}

% ---------------------------------------------------------------------------
% I - INTRODUCTION
% ---------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
\input{introduction}

% ---------------------------------------------------------------------------
% II - PanDA OVERVIEW
% ---------------------------------------------------------------------------
\section{PanDA Overview}
\label{sec:panda_overview}
\input{PandaOverview}

% ---------------------------------------------------------------------------
% III - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
% ---------------------------------------------------------------------------
%\section{Deploying PanDA on a Leadership-scale system}
\section{Deploying PanDA on Titan}
\label{sec:panda_titan}
\input{PandaEvolution}

% ---------------------------------------------------------------------------
% IV - ANALYSIS AND DISCUSSION
% ---------------------------------------------------------------------------
\section{Analysis and Discussion}
\label{sec:panda_titan}
\input{overall}

% ---------------------------------------------------------------------------
% VI - THE NEXT GENERATION EXECUTOR
% ---------------------------------------------------------------------------
% \section{PANDA: Future/RoadMap}
\section{PANDA\@: The Next Generation Executor}
\label{sec:panda_roadmap}
\input{nge}
\input{Experiments}

% ---------------------------------------------------------------------------
% VII - RELATED WORK
% ---------------------------------------------------------------------------
%\vspace{-0.4in}
\section{Related Work}
\label{sec:related}
\input{bckground}

% ---------------------------------------------------------------------------
% VIII - CONCLUSION
% ---------------------------------------------------------------------------
%\vspace{-0.1in}
\section{Conclusion}
\label{sec:conclusion}
\input{conclusion}

% ---------------------------------------------------------------------------
% ACKNOWLEDGEMENTS
% ---------------------------------------------------------------------------
% \section*{Acknowledgements}
% \label{sec:ack}
%
% \input{ack}

% ---------------------------------------------------------------------------
% REFERENCES
% ---------------------------------------------------------------------------
% \newpage
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,bibliography}

% ---------------------------------------------------------------------------
% A - APPENDIX
% ---------------------------------------------------------------------------
% \appendix
% \section{Artifact Description --- Converging High-Throughput and High-Performance Computing: A Case Study}
% \label{apnx:ac}
% \input{appendix}

% that's all folks
\end{document}