### 2017-07-01 PandaEvolution.tex
Address referees' comments and further reductions

% \ifreview
% Comments to address in this section:
% \begin{enumerate}
% 	\item REVIEWER 2: The authors spent quite a few paragraphs for the detail
% 	design and implementation of the PanDA WMS, but introduced few background
% 	of the LHC applications. According to the paper, I understand the
% 	applications running on Titan are a large amount of small independent
% 	jobs. Thus, it seems less cost-effective to run these jobs on a
% 	world-leading supercomputer like Titan which has a fast and expensive
% 	interconnection. It will be more interesting if the authors can compare
% 	the solution on Titan (proposed in this paper) with the original
% 	distributed computing environment and HPECloud from Google mentioned in
% 	the related work section.
% 	\item REVIEWER 5: In Section 3.2, GPUs are not considered in this work,
% 	which would be a waste for using HPC computing. Could this work be
% 	extended to include GPUs or other heterogenous resources?
% \end{enumerate}
% \fi
The upcoming LHC Run 3 will require more resources than the Worldwide LHC
Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more than
300,000 cores at over 100 Grid sites, with a peak performance of 0.3
petaFLOPS\@. This capacity will be sufficient for the planned analysis and
data processing, but it will be insufficient for the Monte Carlo production
workflow and any extra activity. To alleviate these challenges, ATLAS is
expanding the current computing model to include additional resources such as
the opportunistic use of supercomputers.

Generally, supercomputers are designed to support parallel computation that
requires runtime communication. Jobs are executed across multiple cores, each
core calculating a small part of the problem and communicating with other
cores via MPI\@. Accordingly, supercomputers have large number of worker
nodes, connected through a high-speed, low-latency dedicated network. Each
worker node has multicore CPUs, usually augmented with Graphics Processing
Units (GPUs) or other specialized coprocessors.

PanDA WMS has been designed to support distributed Grid computing. Executing
ATLAS workloads or workflows involves concurrent and/or sequential runs of
possibly large amount of jobs, each requiring no or minimal parallelization
and no runtime communication. Thus, computing infrastructure like WLCG have
been designed to aggregate large amount of computing resources across
multiple sites. While each site may deploy MPI capabilities, usually these
are not used to perform distributed computations.

Currently, ATLAS workloads do not require fast interconnects or specilized
coprocessors but supercomputers tend not to reach 100\% utilization due to
the scheduling of jobs requiring large amount of resources. This offers the
possibility to execute ATLAS-like workloads on supercomputers to increase
utilization and reducing the waste of available resources.

We developed a single-point solution to better understand the problem space
of enabling a WMS designed for HTC to execute production workflows on
resources designed to support HPC\@. The PanDA team developed a job broker to
support the execution of part of the ATLAS production Monte Carlo workflow on
Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).

% ---------------------------------------------------------------------------
\subsection{Architectures, Interfaces and Workloads}
\label{ssec:panda-titan}

% The Titan supercomputer, current number three on the Top 500
% list,%~\cite{top500}, 
% is a Cray XK7 system with 18,688 worker nodes and a total
% of 299,008 AMD cores and 18,688 Nvidia Tesla GPUs. 
% Each worker node has an AMD Opteron 6274 16-core CPU,
% a Nvidia Tesla K20X GPU, 32 GB of RAM and no local storage, though a 16 GB
% RAM disk can be set up. Work nodes use Cray’s Gemini interconnect for
% inter-node MPI messaging. Titan is served by the Spider
% II~\cite{oral2013olcf}, a Lustre filesystem with 32 PB of disk storage, and
% by a 29 PB HPSS tape storage system. Titan’s worker nodes run Compute Node
% Linux, a run time environment based on SUSE Linux Enterprise Server.

% Titan's users submit jobs to Titan's PBS scheduler by logging into login or
% data transfer nodes (DTNs). Titan's authentication and authorization model
% is based on two-factor authentication with a RSA SecurID key. Login nodes
% and DTNs have out/inbound wide area network connectivity while worker nodes
% have only local network access. Fair-share and allocation policies are in
% place both for the PBS batch system and shared file systems.

Titan's architecture, configuration and policies poses several challenges to
the integration with PanDA\@. The default deployment model of PanDA Pilot is
unfeasible on Titan: PanDA Pilot is required to contact the Job Dispatcher of
the PanDA Server to pull jobs to execute, but this is not possible on Titan
because worker nodes do not offer outbound network connectivity. Further,
Titan does not support PanDA's security model based on certificates and
virtual organizations, making PanDA's approach to identity management also
unfeasible. While Titan's data transfer nodes (DTNs) offer wide area network
data transfer, an integration with ATLAS DDM is beyond the functional and
administrative scope of the current prototyping phase. Finally, the specific
characteristics of the execution environment, especially the absence of local
storage on the worker nodes and modules tailored to Compute Node Linux,
require re-engineering of ATLAS application frameworks.

Currently, very few HEP applications can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows
can be off-loaded from the Grid to Titan's. Further, when HEP tasks can be
partitioned into independent jobs, Titan worker nodes can be used to execute
up to 16 concurrent payloads, one per each available core. Given these
constraints and challenges, the type of task most suitable for execution at
the moment on Titan is Monte Carlo detector simulation. This type of task is
mostly computational-intensive, requiring less than 2GB of RAM at runtime and
small input data. Detector simulation tasks in ATLAS are performed via
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 detector simulation toolkit~\cite{agostinelli2003geant4}. These tasks
account for $\approx$ 60\% of all the jobs on WLCG, making them a primary
candidate for offloading.

Detector simulation is part of the ATLAS production Monte Carlo (MC)
workflow~\cite{rimoldi2006atlas}.%,de2013delphes,ritsch2014atlas}. 
The MC workflow consists of four main stages: event generation, detector
simulation, digitization, and reconstruction. Event generation creates sets
of particle four-momenta via different generators, e.g.,
PYTHIA,%~\cite{sjostrand2006pythia}, 
HERWIG,%~\cite{corcella2001herwig} 
and many others. Geant4 simulates the ATLAS detector and the interaction
between that detector and particles. Each interaction creates a so-called hit
and all hits are collected and passed on for digitalization, where hits are
further processed to mimic the readout of the detector. Finally,
reconstruction operates local pattern recognition, creating high-level
objects like particles and jets.

% ---------------------------------------------------------------------------
\subsection{PanDA Broker}
\label{ssec:panda_titan}

The lack of wide area network connectivity on Titan's worker nodes is the
most relevant challenge for integrating PanDA WMS and Titan. Without
connectivity, Panda Pilots cannot be scheduled on worker nodes because they
would not be able to communicate with PanDA Server and therefore pull and
execute jobs. This makes impossible to port PanDA Pilot to Titan while
maintaining the defining feature of the pilot abstraction: decoupling
resource acquisition from workload execution via multi-stage scheduling.

The unavailability of pilots is a potential drawback when executing
distributed workloads like MC detector simulation. Pilots are used to
increase the throughput of distributed workloads: while pilots have to wait
in the supercomputer's queue, once scheduled, they can pull and execute jobs
independent from the system's queue. Jobs can be concurrently executed on
every core available to the pilot, and multiple generations of concurrent
executions can be performed until the pilot's walltime is exhausted. This is
particularly relevant for machines like Titan where queue policies privilege
parallel jobs on the base of the number of worker nodes they request: the
higher the number of nodes, the shorter the amount of queue time (modulo
fair-share and allocation policies).

The backfill optimization of Titan's Moab scheduler allows to avoid the
overhead of queue wait times without using pilot
abstraction~\cite{maui_backfill_url}. With this optimization, Moab starts
low-priority jobs when they do not delay higher priority jobs, independent of
whether the low-priority jobs were queued after the high-priority ones.

When the backfill optimization is enabled, users can interrogate Moab about
the number of worker nodes and walltime that would be available to a
low-priority job at that moment in time. If a job is immediately submitted to
Titan with that number of worker nodes and walltime, chances are that Moab
will immediately schedule it, reducing its queue time to a minimum. In this
paper, we call this number of worker nodes and walltime an available
`backfill slot'.

Compared to pilots, backfill has the disadvantage of limiting the amount of
worker nodes that can be requested. Pilots are normal jobs: they can request
as many worker nodes and walltime as a queue can offer. On the contrary, jobs
sized according to an available backfill slot depend on the number of worker
nodes and walltime that cannot be given to any other job at that moment in
time.

At any point in time, the size of an available backfill slot is typically a
small fraction of the total capacity of a resource. Notwithstanding, given
the size of Titan this translates into a substantial capacity. Every year,
about 10\% of Titan's capacity remains unused~\cite{barker2016us},
corresponding to an average of 30,000 unused cores (excluding GPU cores).
This equals to roughly 10\% of the overall capacity of WLCG\@.

Given the communication requirements of PanDA Pilots and the unused capacity
of Titan, PanDA pilot was repurposed to serve as a job broker on the DTN
nodes of Titan (Fig.~\ref{fig:panda_broker}). Maintaining the core modules of
PanDA Pilot and its stand-alone architecture, this prototype called `PanDA
Broker' 
% implements functionalities to: (i) interrogate Titan about backfill
% availability via a dedicated Moab scheduler command; (ii) pull ATLAS jobs
% and events from PanDA Server; (iii) wrap the payload of ATLAS jobs into MPI
% scripts; (iv) submitting workload-specific MPI scripts to Titan's PBS batch
% system and monitor their execution; and (v) staging and preparing
% input/output files. Backfill querying, payload wrapping, and scripts
% submission required a new implementation while pulling ATLAS job and
% events, and file staging were inherited from PanDA Pilot.
%
% Backfill querying is performed via a dedicated Moab scheduler command while
% a tailored Python MPI script is used to execute the payload of ATLAS jobs.
% This enables the execution of unmodified Grid-centric, ATLAS jobs on Titan.
% Typically, a MPI script is workload-specific as it sets up the execution
% environment for a specific payload. This involves organization of worker
% directories, data management, optional input parameters modification, and
% cleanup on exit. Upon submission, a copy of the MPI script runs on every
% available worker node, starting the execution of the ATLAS job's payload in
% a subprocess and waits until its completion.
%
% MPI scripts are submitted to Titan's PBS batch system via
% RADICAL-SAGA~\cite{radical-saga_url}, a Python module, compliant with the
% OGF GFD.90 SAGA specification~\cite{goodale2008simple}. The Simple API for
% Grid Applications (SAGA) offers a unified interface to diverse job
% schedulers and file transferring services. In this way, SAGA provides an
% interoperability layer that lowers the complexity of using distributed
% infrastructures. Behind the API façade, RADICAL-SAGA implements a adaptor
% architecture: each adaptor interface the SAGA API with different middleware
% systems and services, including the PBS batch scheduler of Titan.
%
% The data staging capabilities of the PanDA Broker are implemented via a
% file system shared among DTNs and worker nodes. The input files with the
% events of the ATLAS jobs are downloaded on the shared filesystem from the
% data center of Brookhaven National Laboratory (BNL). The MPI script setup
% process includes making the location of these files available to the
% payload of the ATLAS's jobs. The PanDA Broker can locate the payload's
% output files on the shared filesystem and transfer them from Titan BNL\@.
%
% Once deployed on Titan, every PanDA Broker 
supports the execution of MC detector simulations in 9 steps.

Each PanDA Broker deployed on Titan queries the Job Dispatcher module of the
PanDA server for ATLAS jobs that have been bound to Titan by JEDI
(Fig.~\ref{fig:panda_broker}:1). Upon receiving jobs descriptions, PanDA
Broker pulls jobs' input files from BNL to the OLCF Lustre file system
(Fig.~\ref{fig:panda_broker}:2). PanDA Broker queries Titan's Moab scheduler
about the current available backfill slot (Fig.~\ref{fig:panda_broker}:3) and
creates an MPI script, wrapping enough ATLAS jobs' payload to fit the
backfill slot. PanDA Broker submits the MPI script to the Titan's Torque
batch system via RADICAL-SAGA (Fig.~\ref{fig:panda_broker}:4).

Upon execution on the worker node(s) (Fig.~\ref{fig:panda_broker}:5), the MPI
script initializes and configures the execution environment
(Fig.~\ref{fig:panda_broker}:6), and executes one AthenaMP for each available
work node (Fig.~\ref{fig:panda_broker}:7). AthenaMP retrieves events from
Lustre (Fig.~\ref{fig:panda_broker}:8) and spawns 1 Geant4 event simulation
process on each of the 16 available cores (Fig.~\ref{fig:panda_broker}:9).
Upon completion of each MPI script, PanDA Broker transfer the jobs' output to
BNL (Fig.~\ref{fig:panda_broker}:10), and performs cleanup.

While PanDA Broker implementation is resource specific, it was successfully
ported to other supercomputers, including the HPC2 at the National Research
Center ``Kurchatov Institute'' (NRC-KI)~\cite{belyaev2015integration},
Edison/Cori at the National Energy Research Scientific Computing Center
(NERSC)~\cite{barreiro2016panda}, and SuperMUC at the Leibniz Supercomputing
Centre (LRZ)~\cite{barreiro2016panda}.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
    \vspace{-0.3in}
    \caption{PanDA Broker architecture as deployed on Titan. Numbers
    indicates the execution process of a detector simulation job, part of the
    production ATLAS MC workflow.}
\label{fig:panda_broker}
\end{figure}
\vspace{-0.04in}