### 2017-07-05 introduction.tex
Fix latex and broken reference

\ifreview
Comments to address in this section:
\begin{enumerate}
	\item REVIEWER 1: listing the key observations in one place, and in a
	consistent fashion. I found The 5 points listed in the first part of the
	rebuttal quite compelling, and strongly urge the authors to put them in
	the paper in a prominent location.
	\item REVIEWER 5: I would have preferred to see a more detailed list of
	contributions, including the novelty of the proposed approach as compared
	to prior work.
	\item I am ambivalent about the use of the term ``converging'' in the
	paper's title. While the paper does port workloads to high-performance
	computing resources, it does not deal with compute-oriented optimizations
	(parallelization), lowering communication, etc., which are often major
	concerns in HPC programming. In this sense, the paper is closer in spirit
	to how to port high-throughput apps to high-performance clusters. Agreed?
	{\bf  Need to clarify it is not the task that is being converged but the workload and execution model that is being converged}
	\item REVIEWER 1: A glossary would be welcomed. {\bf Not addressable}
\end{enumerate}
\fi

The Large Hadron Collider (LHC) was created to explore the fundamental
properties of matter. Multiple experiments at LHC have collected and
distributed hundreds of petabytes of data worldwide to hundreds of computer
centers. Thousands of physicists analyze petascale data volumes daily. The
detection of the Higgs Boson in 2013 speaks to the success of the detector
and experiment design, as well as the sophistication of computing systems
devised to analyze the data, which historically, consisted of the federation
of hundreds to thousands of distributed resources, ranging in scale from
small to mid-size resource~\cite{foster2003grid}.

Although the workloads to be executed are comprised of tasks that are
independent of each other, the management of the distribution of workloads
across many heterogeneous resources, the effective utilization of resources
and efficient execution of workloads present non-trivial challenges. Many
software solutions have been developed in response to these challenges. The
CMS experiment, devised a solution based around the
HTCondor~\cite{thain2005distributed} software ecosystem. The
ATLAS~\cite{Aad:2008} experiment utilizes the Production and Distributed
Analysis (PanDA) workload management system~\cite{Maeno2011} (WMS) for
distributed data processing and analysis. The CMS and ATLAS experiments
utilize, arguably the largest production grade distributed computing
solutions and have symbolized the paradigm of {\it high-throughput
computing}, i.e., the effective execution of many independent tasks.

In spite of the impressive scale of the ATLAS distributed computing system,
demand for computing systems will soon significantly outstrip current and
projected supply.   The data volumes that will need analyzing in LHC-Run 3
($~\approx$ 2022) and the high-luminosity era (Run 4) will increase by
factors of 10--100 compared to the current phase (Run 2). There are multiple
levels at which this problem needs to be addressed: the utilization of
emerging parallel architectures (e.g., platforms); algorithmic and advances
in analytical methods (e.g., use of Machine Learning); and the ability to
exploit different platforms (e.g., clouds and supercomputers).

% As the LHC prepares for Run 3 in $~\approx$ 2022 and the high-luminosity
% era (Run 4), it is anticipated that the data volumes that will need
% analyzing will increase by factors of 10-100 compared to the current phase
% (Run 2). Data will be larger in volume but will also require more
% sophisticated computational processing.

This paper represents the experience of how the ATLAS experiment has ``broken
free'' of the traditional computational approach of high-throughput computing
on distributed resources to embrace new platforms, in particular
high-performance computers. Specifically, we discuss the experience of
integrating PanDA WMS with a US DOE leadership machine (Titan) to reach
sustained production scales of approximately 51M core-hours a year.

We demonstrate how Titan is more efficiently utilized by the mixing of small
and short-lived tasks in backfill with regular payloads. Cycles otherwise
unusable (or very difficult to use) are used for science thus, increasing the
overall utilization on Titan without loss of quality-of-service. The
conventional mix of jobs at OLCF cannot be effectively backfilled because of
size, duration, and scheduling policies. Our approach is extensible to any
HPC with ``capability scheduling'' policies.  We also investigate the use of
a pilot-abstraction based task execution runtime system to flexibly execute
ATLAS and other heterogeneous workloads (molecular dynamics) using regular
queues. As such, our approach provides a general solution and investigation
of the convergence of HPC and HTC execution of workloads.

This work demonstrates a viable production route to delivering large amounts
of computing resources to ATLAS and, in the future, to other experimental and
observational use cases.  This broadens the use of leadership computing while
demonstrating how distributed workflows can be integrated with leadership
resources, and effectively accommodating HTC and HPC workloads simultaneously.

This paper also provides: (i) a critical evaluation of the many design and
operational considerations that have been taken to support the sustained,
scalable and production usage of Titan for historically high-throughput
workloads, and (ii) early lessons and guidance as the community looks forward
to designing the next generation of online analytical
platforms~\cite{foap-url}, for how current and future experimental and
observational systems can be integrated with production supercomputers in a
general and extensible manner.