### 2017-07-07 abstract.tex
We have both 51 and 52M core hours.
  Changing 51 to 52M.

The computing systems used by LHC experiments has historically consisted of
the federation of hundreds to thousands of distributed resources, ranging
from small to mid-size resource.  In spite of the impressive scale of the
existing distributed computing solutions, the federation of small to mid-size
resources will be insufficient to meet projected future demands. This paper
is a case study of how the ATLAS experiment has embraced Titan -- a DOE
leadership facility in conjunction with traditional distributed high-
throughput computing to reach sustained production scales of approximately
52M core-hours a years. The three main contributions of this paper are:  (i)
a critical evaluation of design and operational considerations  to support
the sustained, scalable and production usage of Titan;  (ii) a preliminary
characterization of a next generation executor for PanDA to support new
workloads and  advanced execution modes; and (iii) early lessons for how
current and future experimental and observational systems can be integrated
with production supercomputers and other platforms in a general and
extensible manner.