### 2017-08-28 conclusion.tex

\ifreview
Comments to address in this section:
\begin{itemize}
	\color{red} 
	\item REVIEWER 2: How does Summit, the TITAN successor, compare to the
	conclusions? The end of the TITAN life-cycle basically invalidates the
	applicability of all conclusions if this generalisation is missing.
	\item REVIEWER 3: This statement in the "conclusions" section is quite
	exaggerated: "The experience captured in this paper will prove invaluable
	for designing WMS for computational campaigns and will provide a baseline
	to evaluate the relative merits of different approaches." WMS systems for
	LHC experiments have been developed and run in production for many years;
	while this paper adds interesting experience and data points, it's
	unlikely that this will be "invaluable" for new designs and a cornerstone
	for their evaluation.
\end{itemize}
\fi

The deployment of PanDA Broker on Titan enabled distributed computing on a
leadership-class HPC machine at unprecedented scale. In the past 13 months,
PanDA WMS has consumed almost 52M core-hours on Titan, simulating 3.5\% of
the total number of detector events of the ATLAS production Monte Carlo
workflow. We described the implementation and execution process of PanDA WMS
(\S\ref{sec:panda_overview}) and PanDA Broker (\S\ref{sec:panda_titan}),
showing how they support and enable distributed computing at this scale on
Titan, a leadership-class HPC machine managed by OCLF.

We characterized the experience by evaluating the efficiency, scalability and
reliability of both PanDA Broker and AthenaMP as deployed on Titan
(\S\ref{sec:panda_titan}). Our characterization highlighted the strengths and
limitations of the current design and implementation: PanDA Brokers enable
the sustained execution of millions of simulations per week but further work
is required to optimize its efficiency and reliability
(\S\ref{ssec:broker_titan}). PanDA Brokers support the concurrent execution
of multiple AthenaMP instances, enabling each AthenaMP to perform the
concurrent execution of up to 16 Geant4 simulators. Nonetheless, our
characterization showed how improving I/O performance could reduce overheads
(\S\ref{ssec:athenamp_titan}), increasing the overall utilization of Titan's
backfill availability. We introduced PanDA's next generation executor for HPC
systems, characterized its ability to support multi-generation workloads and
analyzed its scaling behavior.  Performance tests on the next generation
executor demonstrate linear strong and weak scalability over several orders
of magnitude  of task and node counts.  Qualitatively, it enables the high-
performance execution of new workloads and advanced execution modes of
traditional workloads.

% More generally, this paper highlights the fundamental role of WMS for
% experimental science.

HEP was amongst the first, if not the first experimental community to realize
the importance of using WMS to manage their computational campaign(s). As
computing becomes increasingly critical for a range of experiments, the
experience foreshadows the importance of WMS for other experiments (such as
SKA, LSST etc.).  These experiments will have their own workload
characteristics, resources types and federation constraints, as well metrics
of performance. The experience captured in this paper will prove invaluable
for designing WMS for computational campaigns and will provide a baseline to
evaluate the relative merits of different approaches.

The 52M core hours used by ATLAS, via PanDA, is over 2\% of the total
utilization on Titan over the same period, bringing the time-averaged
utilization of Titan to be consistently upwards of 90\%. Given that the
average utilization of most other leadership class machines is less (e.g.,
NSF's flagship Blue Waters the average utilization fluctuates between
60--80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
approaches elsewhere. These unprecedented efficiency gains aside, this work
is just a starting point towards more effective operational models for future
leadership and online analytical platforms~\cite{foap-url}. These platforms
will have to support ever increasing complex workloads with varying models
for dynamic resource federation.

% The development of NGE is looking forward to deal with these challenges by
% providing an abstraction layer for HPC resources together with a set of
% APIs for the coordination and fast deployment of heterogeneous workloads.
% In this paper, we provided the results of a preliminary testing of the NGE.
% The testing pointed out the robustness of NGE at handling an high number of
% tasks and worker nodes by mantaining an overhead that grows linearly with
% the number of tasks.

% Recent efforts with Google Cloud for the CMS experiment --
% HEPCloud~\cite{hepcloud,googlehep} represent similar yet different
% approaches to dynamic resource federation. As the type, heterogeneity and
% complexity of future platforms  (e.g., Exascale, Commercial Clouds) for
% science increases, there will be a greater need and emphasis on
% abstractions to ensure WMS can both utilize as well as support evolving
% platforms. Our experience demonstrated the advantages arising from an
% abstractions based approach \textendash{} in particular scalable
% implementations of pilot abstraction.