### 2017-09-10 paper.tex
fixing  title, comments at start.

\documentclass[conference]{IEEEtran}

% ---------------------------------------------------------------------------
% Authors are invited to submit unpublished, original work, using the IEEE
% 8.5 × 11 manuscript guidelines: double-column text using single-spaced 10
% point font on 8.5 × 11 inch pages. Templates are available from http://www.
% ieee.org/conferences_events/conferences/publishing/templates.html.
%
% The conference proceedings will be made available online through the IEEE
% Digital Library.
%
% Three types of papers will be accepted:
%
% - Full research papers (up to 10 pages): focus on new research achievements
%   in eScience, spanning core technologies and applications.
% - Experience papers (up to 10 pages): focus on practical outcomes through
%   applying eScience tools and techniques to novel problems.
% ---------------------------------------------------------------------------

\input{head}
\input{include}

% ---------------------------------------------------------------------------
% BEGIN DOCUMENT
% ---------------------------------------------------------------------------
\begin{document}

\title{High-Throughput Computing on High-Performance Platforms: A Case Study}

%\title{Converging High-Throughput and High-Performance Computing: A Case Study}

\author{
    \IEEEauthorblockN{
        Alessio Angius\IEEEauthorrefmark{1}\IEEEauthorrefmark{2},
        Danila Oleynik\IEEEauthorrefmark{1}\IEEEauthorrefmark{5},
        Sergey Panitkin\IEEEauthorrefmark{1}\IEEEauthorrefmark{4},
        Matteo Turilli\IEEEauthorrefmark{1}\IEEEauthorrefmark{2},\\
        Kaushik De\IEEEauthorrefmark{5},
        Alexei Klimentov\IEEEauthorrefmark{4},
        Sarp Oral\IEEEauthorrefmark{6},
        Jack C. Wells\IEEEauthorrefmark{7} and
        Shantenu Jha\IEEEauthorrefmark{2}
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{1}First authors, alphabetical order
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{2}Department of Electric and Computer Engineering, Rutgers University, NJ, USA
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{4}Physics Department, Brookhaven National Laboratory, NY, USA
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{5}Physics Department, University of Texas Arlington, TX, USA
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{6}Oak Ridge Leadership Computing Facility, Oak Ridge National Laboratory, TN, USA
        }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{7}National Center for Computational Sciences, Oak Ridge National Laboratory, TN, USA
        }
}

% ---------------------------------------------------------------------------
% ABSTRACT
% ---------------------------------------------------------------------------
\maketitle

\ifreview
General comments to address:
\begin{itemize}
    \color{red} 
    % \item REVIEWER 1: There is a problem with capitalization in some
    % reference titles, ie. [2], [3], [4], [7], [8], [9], [10], [11], [12],
    % [13]. Reference [6] is just an author list, year and title. Should there
    % be more here?\mtnote{Fixed.}
    \item REVIEWER 2:  Key concepts such as pilot jobs, tasks and workloads
    are not meaningfully defined.
    % \item REVIEWER 2: the information provided is insufficient to asses in
    % how far any of the presented evaluation, characterization or lessons are
    % applicable outside of TITAN and PanDA Broker combined.
    % \item REVIEWER 2: Several references are of poor quality. For example,
    % [19] is a slide collection even though there are many papers for the
    % referenced work.\mtnote{Reference [19] is a technical paper.}
    % \item REVIEWER 3: Title looks too general for what amounts to the
    % implementation of an adapter component in a system that is very specific
    % to a single HEP experiment. (In general, title, abstract and conclusions
    % seem to over-state the contributions of this paper, considering this is a
    % general eScience conference and not a WLCG-specific one.)
\end{itemize}
\fi

\begin{abstract}
\input{abstract}
\end{abstract}
% \keywords{Distributed computing, Workload management system}

% ---------------------------------------------------------------------------
% I - INTRODUCTION
% ---------------------------------------------------------------------------
\section{Introduction}\label{sec:intro}
\input{introduction}

% ---------------------------------------------------------------------------
% II - PanDA OVERVIEW
% ---------------------------------------------------------------------------
\section{PanDA Overview}\label{sec:panda_overview}
\input{PandaOverview}

% ---------------------------------------------------------------------------
% III - DEPLOYING PANDA ON A LEADERSHIP-SCALE SYSTEM
% ---------------------------------------------------------------------------
\section{Deploying PanDA on Titan}\label{sec:panda_titan}
\input{PandaEvolution}

% ---------------------------------------------------------------------------
% IV - ANALYSIS AND DISCUSSION
% ---------------------------------------------------------------------------
\section{Analysis and Discussion}\label{sec:analysis}
\input{overall}

% ---------------------------------------------------------------------------
% V - THE NEXT GENERATION EXECUTOR
% ---------------------------------------------------------------------------
\section{PANDA\@: The Next Generation Executor}\label{sec:panda_roadmap}
\input{nge}
\input{Experiments}

% ---------------------------------------------------------------------------
% VI - RELATED WORK
% ---------------------------------------------------------------------------
\section{Related Work}\label{sec:related}
\input{bckground}

% ---------------------------------------------------------------------------
% VII - CONCLUSION
% ---------------------------------------------------------------------------
\section{Conclusion}\label{sec:conclusion}
\input{conclusion}

% ---------------------------------------------------------------------------
% REFERENCES
% ---------------------------------------------------------------------------
\bibliographystyle{IEEEtran}
\bibliography{bibliography}

% that's all folks
\end{document}