### 2017-10-28 arxiv/v2/conclusion.tex

The deployment of PanDA Broker on Titan enabled distributed computing on a
leadership-class HPC machine at unprecedented scale. In the past 13 months,
PanDA WMS has consumed almost 52M core-hours on Titan, simulating 3.5\% of
the total number of detector events of the ATLAS production Monte Carlo
workflow. We described the implementation and execution process of PanDA WMS
(\S\ref{sec:panda_overview}) and PanDA Broker (\S\ref{sec:panda_titan}),
showing how they support and enable distributed computing at this scale on
Titan, a leadership-class HPC machine managed by OCLF.

We characterized the efficiency, scalability and reliability of both PanDA
Broker and AthenaMP as deployed on Titan (\S\ref{sec:panda_titan}). Our
characterization highlighted the strengths and limitations of the current
design and implementation: PanDA Brokers enable the sustained execution of
millions of simulations per week but further work is required to optimize its
efficiency and reliability (\S\ref{ssec:broker_titan}). PanDA Brokers support
the concurrent execution of multiple AthenaMP instances, enabling each
AthenaMP to perform the concurrent execution of up to 16 Geant4 simulators.
Nonetheless, our characterization showed how improving I/O performance could
reduce overheads (\S\ref{ssec:athenamp_titan}), increasing the overall
utilization of Titan's backfill availability. We introduced PanDA's next
generation executor for HPC systems, characterized its ability to support
multi-generation workloads and analyzed its scaling behavior.  Performance
tests on the next generation executor demonstrate linear strong and weak
scalability over several orders of magnitude  of task and node counts.
Qualitatively, it enables the high-performance execution of new workloads and
advanced execution modes of traditional workloads.

HEP was amongst the first, if not the first experimental community to realize
the importance of using WMS to manage their computational campaign(s). As
computing becomes increasingly critical for a range of experiments, the
experience foreshadows the importance of WMS for other experiments (such as
SKA, LSST etc.).  These experiments will have their own workload
characteristics, resources types and federation constraints, as well metrics
of performance. The experience captured in this paper should be useful for
designing WMS for computational campaigns and will provide a baseline to
evaluate the relative merits of different approaches.

The 52M core hours used by ATLAS, via PanDA, is over 2\% of the total
utilization on Titan over the same period, bringing the time-averaged
utilization of Titan to be consistently upwards of 90\%. Given that the
average utilization of most other leadership class machines is less (e.g.,
NSF's flagship Blue Waters the average utilization fluctuates between
60--80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
approaches elsewhere. These unprecedented efficiency gains aside, this work
is just a starting point towards more effective operational models for future
leadership and online analytical platforms~\cite{foap-url}. These platforms
will have to support ever increasing complex workloads with varying models
for dynamic resource federation.

{\bf \footnotesize Acknowledgements: This work is funded by Award number
DE-SC0016280 from the Office of Advanced Scientific Computing Research within
the Department of Energy. This research used resources of the Oak Ridge
Leadership Computing Facility at the Oak Ridge National Laboratory, which is
supported by the Office of Science of the U.S. Department of Energy under
Contract No. DE-AC05-00OR22725. \par}